---
title: "yes_data_binding"
author: "Mallory Dobias"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    toc: yes
    toc_depth: 5
    toc_float:
      collapsed: no
      smooth_scroll: no
  pdf_document:
    toc: yes
    toc_depth: '5'
  word_document:
    toc: yes
    toc_depth: '5'
geometry: margin=0.50in
---

```{r setup, include=FALSE}

## Setting up R

knitr::opts_chunk$set(echo = TRUE)

if(!require(pointblank)){install.packages('pointblank')}
library(pointblank) # installing and calling pointblank package early in Markdown

validate_rmd(summary = TRUE, log_to_file = FALSE) # allowing validation from pointblank package within Markdown

```

```{r, include=FALSE}

if(!require(here)){install.packages('here')}
library(here)

if(!require(dplyr)){install.packages('dplyr')}
library(dplyr)

if(!require(stringr)){install.packages('stringr')}
library(stringr)

if(!require(tidyr)){install.packages('tidyr')}
library(tidyr)

if(!require(readr)){install.packages('readr')}
library(readr)

if(!require(purrr)){install.packages('purrr')}
library(purrr)

```

# Data Reading

```{r reading in data}

## Reading in all cleaned english survey data

library(here) # places working directory at root of project folder

d01_data <- readRDS(here("data", "clean", "yes_data_d01_english_clean.Rds")) # directing to data from project folder root

d02_data <- readRDS(here("data", "clean", "yes_data_d02_english_clean.Rds"))

d03_data <- readRDS(here("data", "clean", "yes_data_d03_english_clean.Rds")) 

d05_data <- readRDS(here("data", "clean", "yes_data_d05_english_clean.Rds")) 

d07_data <- readRDS(here("data", "clean", "yes_data_d07_turkish_clean.Rds")) 

d09_data <- readRDS(here("data", "clean", "yes_data_d09_english_clean.Rds")) 

d16_data <- readRDS(here("data", "clean", "yes_data_d16_english_clean.Rds")) 

```

# Binding Data

```{r binding data}

## Binding data from all english sources

english_data_bound <- bind_rows(d02_data, d03_data, d05_data, d09_data, d16_data, d01_data) %>%
  mutate(data_id = factor(data_id, levels = c("d01", "d02", "d03", "d05", "d09", "d16"))) %>%
  arrange(data_id)

## Binding data from all sources (across languages)

all_data_bound <- bind_rows(d02_data, d03_data, d05_data, d07_data, d09_data, d16_data, d01_data) %>%
  mutate(data_id = factor(data_id, levels = c("d01", "d02", "d03", "d05", "d07", "d09", "d16"))) %>%
  arrange(data_id)

```

Checking final English data has expected number of rows

```{r checking n rows english data, validate = TRUE}

## Checking if the row count of data equals the sum of row counts of component dfs

validation_result <- 
  english_data_bound %>%
  expect_row_count_match(nrow(d01_data) + nrow(d02_data) + nrow(d03_data) + nrow(d05_data) + nrow(d09_data) + nrow(d16_data))

## Printing the validation result - runs if passes

validation_result

```

```{r checking n rows all data, validate = TRUE}

## Checking if the row count of data equals the sum of row counts of component dfs

validation_result_all <- 
  all_data_bound %>%
  expect_row_count_match(nrow(d01_data) + nrow(d02_data) + nrow(d03_data) + 
                         nrow(d07_data) + nrow(d05_data) + nrow(d09_data) + nrow(d16_data))

## Printing the validation result - runs if passes

validation_result_all

```

Checking final data has expected columns

```{r pulling colnames for all dfs}

## Creating dfs of col names from data and each component df

data_cols_english <- data.frame(column_name = names(english_data_bound))

data_cols_all <- data.frame(column_name = names(all_data_bound))

d01_cols <- data.frame(column_name = names(d01_data))

d02_cols <- data.frame(column_name = names(d02_data))

d03_cols <- data.frame(column_name = names(d03_data))

d05_cols <- data.frame(column_name = names(d05_data))

d07_cols <- data.frame(column_name = names(d07_data))

d09_cols <- data.frame(column_name = names(d09_data))

d16_cols <- data.frame(column_name = names(d16_data))

```

```{r confirming no new cols in english data, message=FALSE}

## Confirming no columns exist in final df that are not in a component df

n_new_cols <- data_cols_english %>%
  anti_join(d01_cols) %>%
  anti_join(d02_cols) %>%
  anti_join(d03_cols) %>%
  anti_join(d05_cols) %>%
  anti_join(d09_cols) %>%
  anti_join(d16_cols) %>%
  nrow()

pass_new_cols <- n_new_cols == 0

pass_new_cols # true if passes and no new cols

```

```{r confirming no new cols in all data, message=FALSE}

## Confirming no columns exist in final df that are not in a component df

n_new_cols <- data_cols_all %>%
  anti_join(d01_cols) %>%
  anti_join(d02_cols) %>%
  anti_join(d03_cols) %>%
  anti_join(d05_cols) %>%
  anti_join(d07_cols) %>%
  anti_join(d09_cols) %>%
  anti_join(d16_cols) %>%
  nrow()

pass_new_cols <- n_new_cols == 0

pass_new_cols # true if passes and no new cols

```

```{r calculating expected v actual n cols for d02}

## Pulling expected vs actual cols from d02 df

n_expected_cols_d02 <- d02_cols # colnames from original df

n_actual_cols_d02_english <- data_cols_english %>%
  inner_join(d02_cols) # colnames in bound df

n_actual_cols_d02_all <- data_cols_all %>%
  inner_join(d02_cols) # colnames in bound df

```

Checking final data has expected number of columns in English data

```{r checking n columns d02 english, validate = TRUE}

## Checking if all colnames from d02 are also in final data

## number of colnames should match between original df and those pulled from 
## the final dataset

validation_result <- 
  n_expected_cols_d02 %>%
  expect_row_count_match(nrow(n_actual_cols_d02_english))

## Printing the validation result - runs if passes

validation_result

```

Checking final data has expected number of columns in all data

```{r checking n columns d02 all, validate = TRUE}

## Checking if all colnames from d02 are also in final data

## number of colnames should match between original df and those pulled from 
## the final dataset

validation_result_all <- 
  n_expected_cols_d02 %>%
  expect_row_count_match(nrow(n_actual_cols_d02_all))

## Printing the validation result - runs if passes

validation_result_all

```

```{r custom function to check colnames}

# Custom function to calculate expected and actual column counts, and validate them
validate_column_counts <- function(original_cols, bound_cols) {
  
  # Step 1: Calculate the expected column count (colnames from the original dataframe)
  n_expected_cols <- original_cols
  
  # Step 2: Calculate the actual column count (matching colnames in the bound dataframe)
  n_actual_cols <- bound_cols %>%
    inner_join(original_cols, by = "column_name")
  
  # Step 3: Validation check if the actual column count matches the expected column count
  # validation_result <- expect_equal(n_actual_cols, n_expected_cols, 
  #                                   label = "Column count match validation")
  
  validation_result <- 
    n_expected_cols %>%
    expect_row_count_match(nrow(n_actual_cols))
  
  # Return the validation result and the column counts
  return(list(
    n_expected_cols = n_expected_cols,
    n_actual_cols = n_actual_cols,
    validation_result = validation_result
  ))
}

```

```{r d01 validation, validate=TRUE}

## Running the validation function

validation_result_d01 <- validate_column_counts(d01_cols, data_cols_english)

validation_result_d01_all <- validate_column_counts(d01_cols, data_cols_all)

## Printing results

print(validation_result_d01$validation_result) # runs if passes

print(validation_result_d01_all$validation_result) # runs if passes

```

```{r d02 validation, validate=TRUE}

## Running the validation function

validation_result_d02 <- validate_column_counts(d02_cols, data_cols_english)

validation_result_d02_all <- validate_column_counts(d02_cols, data_cols_all)

## Printing results

print(validation_result_d02$validation_result) # runs if passes

print(validation_result_d02_all$validation_result) # runs if passes

```

```{r d03 validation, validate=TRUE}

## Running the validation function

validation_result_d03 <- validate_column_counts(d03_cols, data_cols_english)

validation_result_d03_all <- validate_column_counts(d03_cols, data_cols_all)

## Printing results

print(validation_result_d03$validation_result) # runs if passes

print(validation_result_d03_all$validation_result) # runs if passes

```

```{r d05 validation, validate=TRUE}

## Running the validation function

validation_result_d05 <- validate_column_counts(d05_cols, data_cols_english)

validation_result_d05_all <- validate_column_counts(d05_cols, data_cols_all)

## Printing results

print(validation_result_d05$validation_result) # runs if passes

print(validation_result_d05_all$validation_result) # runs if passes

```

```{r d07 validation, validate=TRUE}

## Running the validation function

validation_result_d07_all <- validate_column_counts(d07_cols, data_cols_all) # d07 is not in the english df (turkish)

## Printing results

print(validation_result_d07_all$validation_result) # runs if passes

```

```{r d09 validation, validate=TRUE}

## Running the validation function

validation_result_d09 <- validate_column_counts(d09_cols, data_cols_english)

validation_result_d09_all <- validate_column_counts(d09_cols, data_cols_all)

## Printing results

print(validation_result_d09$validation_result) # runs if passes

print(validation_result_d09_all$validation_result) # runs if passes

```

```{r d16 validation, validate=TRUE}

## Running the validation function

validation_result_d16 <- validate_column_counts(d16_cols, data_cols_english)

validation_result_d16_all <- validate_column_counts(d16_cols, data_cols_all)

## Printing results

print(validation_result_d16$validation_result) # runs if passes

print(validation_result_d16_all$validation_result) # runs if passes

```

Visual inspection of unique cols to each df

```{r pulling unique cols, message=FALSE}

## Pulling cols unique to each df 

## visual inspection here will help find vars that need renaming
## to keep consistency between dfs var naming

unique_cols_d01 <- d01_cols %>%
  anti_join(d02_cols) %>%
  anti_join(d03_cols) %>%
  anti_join(d05_cols) %>%
  anti_join(d07_cols) %>%
  anti_join(d09_cols) %>%
  anti_join(d16_cols)

unique_cols_d02 <- d02_cols %>%
  anti_join(d01_cols) %>%
  anti_join(d03_cols) %>%
  anti_join(d05_cols) %>%
  anti_join(d07_cols) %>%
  anti_join(d09_cols) %>%
  anti_join(d16_cols)

unique_cols_d03 <- d03_cols %>%
  anti_join(d01_cols) %>%
  anti_join(d02_cols) %>%
  anti_join(d05_cols) %>%
  anti_join(d07_cols) %>%
  anti_join(d09_cols) %>%
  anti_join(d16_cols)

unique_cols_d05 <- d05_cols %>%
  anti_join(d01_cols) %>%
  anti_join(d02_cols) %>%
  anti_join(d03_cols) %>%
  anti_join(d07_cols) %>%
  anti_join(d09_cols) %>%
  anti_join(d16_cols)

unique_cols_d07 <- d07_cols %>%
  anti_join(d01_cols) %>%
  anti_join(d02_cols) %>%
  anti_join(d03_cols) %>%
  anti_join(d05_cols) %>%
  anti_join(d09_cols) %>%
  anti_join(d16_cols)

unique_cols_d09 <- d09_cols %>%
  anti_join(d01_cols) %>%
  anti_join(d02_cols) %>%
  anti_join(d03_cols) %>%
  anti_join(d05_cols) %>%
  anti_join(d07_cols) %>%
  anti_join(d16_cols)

unique_cols_d16 <- d16_cols %>%
  anti_join(d01_cols) %>%
  anti_join(d02_cols) %>%
  anti_join(d03_cols) %>%
  anti_join(d05_cols) %>%
  anti_join(d07_cols) %>%
  anti_join(d09_cols)

```

```{r printing unique cols}

## Printing colnames unique to each df

unique_cols_d01 %>%
  rename(d01_column_name = column_name)

unique_cols_d02 %>%
  rename(d02_column_name = column_name)

unique_cols_d03 %>%
  rename(d03_column_name = column_name)

unique_cols_d05 %>%
  rename(d05_column_name = column_name)

unique_cols_d07 %>%
  rename(d07_column_name = column_name)

unique_cols_d09 %>%
  rename(d09_column_name = column_name)

unique_cols_d16 %>%
  rename(d16_column_name = column_name)

```

Reorganizing column order in english data

```{r reorganizing cols english data}

english_data_reorged <- english_data_bound %>%
  select(response_id, language, data_id, contains("date"),
         perc_progress, finished, ssi_choice, ssi_source,
         starts_with("volunteer_") & !contains("tim"),
         contains("reliant"), contains("san_antonio"), contains("middle_east_"),
         contains("hear_proj"), dem_why_interest, dem_receive_support, dem_why_no_support, dem_location, dem_adult, dem_age, dem_sex,
         starts_with("dem_gender_") & !contains("tim"),
         starts_with("dem_sex_orientation") & !contains("tim"), 
         starts_with("dem_race") & !contains("tim"),
         starts_with("dem_ethnicity") & !contains("tim"),
         starts_with("ssi_prev"), starts_with("ssi_times"), 
         starts_with("mfq_") & !contains("tim"), 
         starts_with("pre_bhs_") & !contains("tim"), 
         starts_with("pre_shs_") & !contains("tim"), 
         starts_with("pre_self_hate_") & !contains("tim"),
         starts_with("pre_perceived_control") & !contains("tim"),
         ssi_started,
         starts_with("abc") & !contains("tim"),
         starts_with("proj_care") & !contains("tim"),
         starts_with("proj_pers") & !contains("tim"),
         starts_with("proj_rise") & !contains("tim"),
         ssi_completed,
         starts_with("advice") & !contains("tim"),
         starts_with("pfs_") & !contains("tim"),
         starts_with("perc_change_") & !contains("tim"),
         starts_with("post_bhs_") & !contains("tim"),
         starts_with("post_shs_") & !contains("tim"),
         starts_with("post_self_hate_") & !contains("tim"),
         starts_with("post_perceived_control") & !contains("tim"),
         contains("tim") & starts_with("select_language"),
         contains("tim") & starts_with("intro"),
         contains("tim") & starts_with("into"),
         contains("tim") & starts_with("volunteer"),
         contains("tim") & starts_with("dem"),
         contains("tim") & starts_with("ssi"),
         contains("tim") & starts_with("mfq"),
         contains("tim") & starts_with("pre_"),
         contains("tim") & starts_with("abc"),
         contains("tim") & starts_with("proj_care"),
         contains("tim") & starts_with("proj_rise"),
         contains("tim") & starts_with("proj_pers"),
         contains("tim") & starts_with("advice_"),
         contains("tim") & starts_with("pfs_"),
         contains("tim") & starts_with("perc_change_"),
         contains("tim") & starts_with("post_"))

```

```{r checking col order setup english data, include=FALSE}

## creating data frames of names pre and post reorg for comparison

pre_reorg_names <- english_data_bound %>%
  names() # names pre

post_reorg_names <- english_data_reorged %>%
  names() # names post

# Create tibbles for easier manipulation

pre_reorg_df <- tibble(names = pre_reorg_names)
post_reorg_df <- tibble(names = post_reorg_names)

# Get the unique names (names not in both data frames)

test_reorg <- bind_rows(pre_reorg_df, post_reorg_df) %>%
  group_by(names) %>%
  filter(n() == 1)
         
```

Checking that reorganizing happened correctly (i.e., no columns dropped) in english data

```{r checking reorg did not lose cols 2 english data}

## Checking if df with any columns unique to pre-reorg data is empty (i.e., no columns dropped)

n_cols_dropped <- test_reorg %>%
  nrow() == 0

n_cols_dropped # true if no columns were dropped by reorg

```

Reorganizing column order in turkish data (no binding necessary)

```{r reorganizing cols turkish data}

turkish_data_reorged <- d07_data %>%
  select(response_id, language, data_id, contains("date"),
         perc_progress, finished, q_recaptcha_score, ssi_choice,
         starts_with("volunteer_") & !contains("tim"),
         contains("reliant"), contains("san_antonio"), contains("middle_east_"),
         contains("hear_proj"), dem_why_interest, dem_receive_support, dem_why_no_support, dem_age, dem_sex,
         starts_with("dem_race") & !contains("tim"),
         starts_with("dem_ethnicity") & !contains("tim"),
         starts_with("ssi_prev"), starts_with("ssi_times"), 
         starts_with("mfq_") & !contains("tim"), 
         starts_with("pre_bhs_") & !contains("tim"), 
         starts_with("pre_shs_") & !contains("tim"), 
         starts_with("pre_self_hate_") & !contains("tim"),
         starts_with("pre_perceived_control") & !contains("tim"),
         ssi_started,
         starts_with("abc") & !contains("tim"),
         starts_with("proj_care") & !contains("tim"),
         starts_with("proj_pers") & !contains("tim"),
         starts_with("proj_rise") & !contains("tim"),
         ssi_completed,
         starts_with("advice") & !contains("tim"),
         starts_with("pfs_") & !contains("tim"),
         starts_with("perc_change_") & !contains("tim"),
         starts_with("post_bhs_") & !contains("tim"),
         starts_with("post_shs_") & !contains("tim"),
         starts_with("post_self_hate_") & !contains("tim"),
         starts_with("post_perceived_control") & !contains("tim"),
         contains("tim") & starts_with("select_language"),
         contains("tim") & starts_with("intro"),
         contains("tim") & starts_with("into"),
         contains("tim") & starts_with("volunteer"),
         contains("tim") & starts_with("dem"),
         contains("tim") & starts_with("ssi"),
         contains("tim") & starts_with("mfq"),
         contains("tim") & starts_with("pre_"),
         contains("tim") & starts_with("abc"),
         contains("tim") & starts_with("proj_care"),
         contains("tim") & starts_with("proj_rise"),
         contains("tim") & starts_with("proj_pers"),
         contains("tim") & starts_with("advice_"),
         contains("tim") & starts_with("pfs_"),
         contains("tim") & starts_with("perc_change_"),
         contains("tim") & starts_with("post_"))

```

```{r checking col order setup turkish data, include=FALSE}

## creating data frames of names pre and post reorg for comparison

pre_reorg_names <- d07_data %>%
  names() # names pre

post_reorg_names <- turkish_data_reorged %>%
  names() # names post

# Create tibbles for easier manipulation

pre_reorg_df <- tibble(names = pre_reorg_names)
post_reorg_df <- tibble(names = post_reorg_names)

# Get the unique names (names not in both data frames)

test_reorg <- bind_rows(pre_reorg_df, post_reorg_df) %>%
  group_by(names) %>%
  filter(n() == 1)
         
```

Checking that reorganizing happened correctly (i.e., no columns dropped) in turkish data

```{r checking reorg did not lose cols 2 turkish data}

## Checking if df with any columns unique to pre-reorg data is empty (i.e., no columns dropped)

n_cols_dropped <- test_reorg %>%
  nrow() == 0

n_cols_dropped # true if no columns were dropped by reorg

```

Reorganizing cols in all data df

```{r reorganizing cols all data}

all_data_reorged <- all_data_bound %>%
  select(response_id, language, data_id, contains("date"),
         perc_progress, finished, q_recaptcha_score, ssi_choice, ssi_source,
         starts_with("volunteer_") & !contains("tim"),
         contains("reliant"), contains("san_antonio"), contains("middle_east_"),
         contains("hear_proj"), dem_why_interest, dem_receive_support, dem_why_no_support, dem_location, dem_adult, dem_age, dem_sex,
         starts_with("dem_gender_") & !contains("tim"),
         starts_with("dem_sex_orientation") & !contains("tim"), 
         starts_with("dem_race") & !contains("tim"),
         starts_with("dem_ethnicity") & !contains("tim"),
         starts_with("ssi_prev"), starts_with("ssi_times"), 
         starts_with("mfq_") & !contains("tim"), 
         starts_with("pre_bhs_") & !contains("tim"), 
         starts_with("pre_shs_") & !contains("tim"), 
         starts_with("pre_self_hate_") & !contains("tim"),
         starts_with("pre_perceived_control") & !contains("tim"),
         ssi_started,
         starts_with("abc") & !contains("tim"),
         starts_with("proj_care") & !contains("tim"),
         starts_with("proj_pers") & !contains("tim"),
         starts_with("proj_rise") & !contains("tim"),
         ssi_completed,
         starts_with("advice") & !contains("tim"),
         starts_with("pfs_") & !contains("tim"),
         starts_with("perc_change_") & !contains("tim"),
         starts_with("post_bhs_") & !contains("tim"),
         starts_with("post_shs_") & !contains("tim"),
         starts_with("post_self_hate_") & !contains("tim"),
         starts_with("post_perceived_control") & !contains("tim"),
         contains("tim") & starts_with("select_language"),
         contains("tim") & starts_with("intro"),
         contains("tim") & starts_with("into"),
         contains("tim") & starts_with("volunteer"),
         contains("tim") & starts_with("dem"),
         contains("tim") & starts_with("ssi"),
         contains("tim") & starts_with("mfq"),
         contains("tim") & starts_with("pre_"),
         contains("tim") & starts_with("abc"),
         contains("tim") & starts_with("proj_care"),
         contains("tim") & starts_with("proj_rise"),
         contains("tim") & starts_with("proj_pers"),
         contains("tim") & starts_with("advice_"),
         contains("tim") & starts_with("pfs_"),
         contains("tim") & starts_with("perc_change_"),
         contains("tim") & starts_with("post_"))

```

```{r checking col order setup all data, include=FALSE}

## creating data frames of names pre and post reorg for comparison

pre_reorg_names <- all_data_bound %>%
  names() # names pre

post_reorg_names <- all_data_reorged %>%
  names() # names post

# Create tibbles for easier manipulation

pre_reorg_df <- tibble(names = pre_reorg_names)
post_reorg_df <- tibble(names = post_reorg_names)

# Get the unique names (names not in both data frames)

test_reorg <- bind_rows(pre_reorg_df, post_reorg_df) %>%
  group_by(names) %>%
  filter(n() == 1)
         
```

Checking that reorganizing happened correctly (i.e., no columns dropped) in all data

```{r checking reorg did not lose cols 2 all data}

## Checking if df with any columns unique to pre-reorg data is empty (i.e., no columns dropped)

n_cols_dropped <- test_reorg %>%
  nrow() == 0

n_cols_dropped # true if no columns were dropped by reorg

```

# Saving Data

```{r saving cleaned bound english data}

english_data <- english_data_reorged

# Define the folder and file path

folder_path <- here("data", "clean")  # Path to my desired folder
file_name <- "yes_data_english_clean.Rds"
full_path <- file.path(folder_path, file_name)

# Save the dataframe to the .csv file

saveRDS(english_data, full_path)

```

```{r saving cleaned turkish data}

turkish_data <- turkish_data_reorged

# Define the folder and file path

folder_path <- here("data", "clean")  # Path to my desired folder
file_name <- "yes_data_turkish_clean.Rds"
full_path <- file.path(folder_path, file_name)

# Save the dataframe to the .csv file

saveRDS(turkish_data, full_path)

```

```{r saving cleaned bound all data}

all_data <- all_data_reorged

# Define the folder and file path

folder_path <- here("data", "clean")  # Path to my desired folder
file_name <- "yes_data_all_languages_clean.Rds"
full_path <- file.path(folder_path, file_name)

# Save the dataframe to the .csv file

saveRDS(all_data, full_path)

```