---
title: "yes_data_cleaning"
author: "Mallory Dobias"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    toc: yes
    toc_depth: 5
    toc_float:
      collapsed: no
      smooth_scroll: no
  pdf_document:
    toc: yes
    toc_depth: '5'
  word_document:
    toc: yes
    toc_depth: '5'
geometry: margin=0.50in
---

```{r setup, include=FALSE}

## Setting up R

knitr::opts_chunk$set(echo = TRUE)

if(!require(pointblank)){install.packages('pointblank')}
library(pointblank) # installing and calling pointblank package early in Markdown

validate_rmd(summary = TRUE, log_to_file = FALSE) # allowing validation from pointblank package within Markdown

```

```{r, include=FALSE}

if(!require(here)){install.packages('here')}
library(here)

if(!require(janitor)){install.packages('janitor')}
library(janitor)

if(!require(dplyr)){install.packages('dplyr')}
library(dplyr)

if(!require(stringr)){install.packages('stringr')}
library(stringr)

if(!require(pointblank)){install.packages('pointblank')}
library(pointblank)

```

# Data Reading

```{r reading in data}

## Reading in NU English survey data

library(here) # places working directory at root of project folder

raw_data <- read.csv(here("data", "raw", "NU - YES - English - Questionnaires Optional_October 31, 2024_14.42.csv")) # directing to data from project folder root

# head(raw_data)

```

# Data Cleaning

## Cleaning Var Names

Removing Qualtrics metadata rows

```{r removing first two Qualtrics metadata rows}

## Removing first two rows with Qualtrics metadata

data_sliced <- raw_data %>%
  dplyr::slice(-1,-2)

# head(data_sliced) # looks good

```

```{r checking metadata removal setup, include = FALSE}

## checking Qualtrics metadata rows are removed

# pulling row 3 of raw data to compare to row 1 of sliced data

test_sliced_pre <- raw_data %>%
  dplyr::slice(3) # pulling pre-removal row 3 for comparison

test_sliced_post <- data_sliced %>%
  dplyr::slice(1) # pulling post-removal row 1 for comparison

# combining into one dataframe

test_sliced <- rbind(test_sliced_pre, test_sliced_post)

```

Checking metadata rows removed correctly

```{r checking metadata removal, validate = TRUE}

test_sliced %>%
  pointblank::col_vals_equal(vars(ResponseId), 
                             vars(ResponseId)) # passes if third row response is now first row

```

Converting to tidy var names

```{r creating tidy names}

## Cleaning var names

data_named <- data_sliced %>%
  janitor::clean_names()

# head(data_named)


```

Checking var names are tidy

```{r checking tidy names setup}

## checking var names are now tidy

test_var_names <- tibble(col_names = names(data_named)) # creates data frame of names

test_var_names %>%
  filter(stringr::str_detect(col_names, "[A-Z]")) %>%
  nrow() == 0 # true if zero col names contain capital letters

```

Renaming unlabeled project rise vars

```{r naming rise vars}

## Naming RISE vars

# first insert an underscore, then replace q vars with rise var name labels

data_rise_named <- data_named %>%
  rename_with(~str_replace(., "^q([0-9]+)", "q_\\1"), starts_with("q")) %>%
  rename_with(~ str_replace(., "^q_", "proj_rise_q_"), starts_with("q_"))

```

```{r checking rise var naming setup, include = FALSE}

## checking rise vars are named correctly

test_rise_pre <- data_named %>%
  select(starts_with("q"), -questionnaires, -contains("text")) %>%
  rename_with(~ str_extract(., "\\d+")) %>% # extract digits from rise var names pre cleaning
  colnames() # create df of pre-named var names for comparison

test_rise_post <- data_rise_named %>%
  select(starts_with("proj_rise_"), -contains("commit"), -contains("text")) %>%
  rename_with(~ str_extract(., "\\d+")) %>% # extract digits from rise var names post cleaning
  colnames() # create df of post-named var names for comparison

test_rise_naming_df <- tibble(test_rise_pre, test_rise_post) # creating df for validation

```

Checking columns renamed correctly

```{r checking var naming, validate = TRUE}

test_rise_naming_df %>%
  pointblank::col_vals_equal(vars(test_rise_pre), 
                             vars(test_rise_post)) # passes if renamed columns are identical to original cols

```

## Calculating Composites

Calculating depressive symptoms using the Short Form Mood and Feelings Questionnaire (only measured pre-SSI). Scores range from 0 to 26, where 0 = not true, 1 = sometimes, 2 = true. Scores of 12 or higher suggest depression. Source: https://doi.org/10.1002/mpr.1610

```{r pulling mfq values, include=FALSE}

## Pulling mfq variables

# vars_mfq <- data_rise_named %>%
#   select(contains("mfq")) %>%
#   head() %>%
#   print()

```

Mutating mfq character values to numeric values

```{r mutating mfq values}

## Mutating character values to numeric across multiple mfq columns 

data_mfq_mutated <- data_rise_named %>%
  mutate(across(starts_with("mfq_") & -contains("tim"), 
                ~ case_when(
                    . == "Not True" ~ 0,
                    . == "Sometimes" ~ 1,
                    . == "True" ~ 2,
                    TRUE ~ NA_real_ # Handle other cases with NA
                )))

```

Checking mfq variables are now numeric class

```{r mfq numeric class validation, validate=TRUE}

data_mfq_mutated %>%
  pointblank::col_is_numeric(vars(mfq_1)) # true if mfq vars are now numeric

```

```{r checking mfq mutation setup, include=FALSE}

## checking mutation happened correctly - setting up for pointblank tests

# creating df of count for each response, pre mutation

test_mfq_pre <- data_rise_named %>%
  group_by(mfq_1) %>% # grouping by response
  summarize(count = n()) %>% # counting responses pre-mutation
  arrange(desc(count)) %>% # arranged in descending count order
  rename(mfq_1_pre = mfq_1, count_pre = count) # renaming before combining to one df

# creating df of count for each response, post mutation

test_mfq_post <- data_mfq_mutated %>%
  mutate(across(starts_with("mfq_"), 
                ~ factor(., levels = unique(.)))) %>% # converting to factor class so we can count
  group_by(mfq_1) %>% # grouping by response
  summarize(count = n()) %>% # counting responses post-mutation
  arrange(desc(count)) %>% # arranged in descending count order
  rename(mfq_1_post = mfq_1, count_post = count) # renaming before combining to one df

# creating overall df for comparison

test_mfq_mutate <- cbind(test_mfq_pre, test_mfq_post)

```

```{r writing a function to setup future mutation checks, include=FALSE}

# Custom function to compare value counts before and after mutation
prep_mutation_check <- function(data_pre, data_post, var) {
  
  # Step 1: Create pre-mutation counts
  test_pre <- data_pre %>%
    select(var) %>%
    group_by(across(all_of(var))) %>%  # Group by the specified variable
    summarize(count = n(), .groups = "drop") %>%  # Count occurrences
    arrange(desc(count)) %>%  # Arrange in descending order
    rename_with(~ paste0(., "_pre"), everything())  # Rename columns for pre-mutation
  
  # Step 2: Create post-mutation counts
  test_post <- data_post %>%
    select(var) %>%
    mutate(across(all_of(var), ~ factor(., levels = unique(.)))) %>%  # Convert to factor
    group_by(across(all_of(var))) %>%  # Group by the specified variable
    summarize(count = n(), .groups = "drop") %>%  # Count occurrences
    arrange(desc(count)) %>%  # Arrange in descending order
    rename_with(~ paste0(., "_post"), everything())  # Rename columns for post-mutation
  
  # Step 3: Combine pre and post counts into one dataframe for comparison
  test_comparison <- bind_cols(test_pre, test_post)
  
  return(test_comparison)
}

```

```{r testing function to setup future mutation checks, include=FALSE}

# Example usage:
# Assuming `data_rise_named` is the pre-mutation dataframe
# and `data_mfq_mutated` is the post-mutation dataframe,
# and you want to compare the value counts of "mfq_1":

test_mutate <- prep_mutation_check(data_rise_named, data_mfq_mutated, "mfq_1") # same output as test_mfq_mutate above

## will use the prep_mutation_check() function to set up mutation checks moving forward

```

Checking mfq response counts are identical from pre to post mutation

```{r mfq response counts validation, validate = TRUE}

test_mutate %>%
  pointblank::col_vals_equal(vars(count_pre), 
                             vars(count_post)) # true if response counts are identical pre to post mutation

```

Calculating mfq composite scores

```{r calculating mfq scores}

## Calculating composite mfq scores

data_mfq_scored <- data_mfq_mutated %>%
  mutate(mfq_total = rowSums(across(starts_with("mfq_") & -contains("tim"))))

```

```{r checking mfq score setup, include = FALSE}

## checking composite calculation was correct

# manually calculating mfq totals to compare sum methods

test_mfq_calc <- data_mfq_scored %>%
  select(response_id, starts_with("mfq_") & -contains("tim")) %>%
  mutate(manual_mfq_total = mfq_1 + mfq_2 + mfq_3 + mfq_4 + mfq_5 +
                            mfq_6 + mfq_7 + mfq_8 + mfq_9 + mfq_10 +
                            mfq_11 + mfq_12 + mfq_13) %>%
  select(mfq_total, manual_mfq_total)

```

Checking mfq totals are identical across sum methods

```{r mfq scoring validation, validate = TRUE}

test_mfq_calc %>%
  pointblank::col_vals_equal(vars(mfq_total), 
                             vars(manual_mfq_total), 
                             na_pass = TRUE) # true if mfq totals are identical across sum methods

```

Calculating hopelessness using the 4-item Beck's Hopelessness Scale (measured pre and post SSI). Mean scores range from 0 to 3, where 0 = absolutely disagree, 1 = somewhat disagree, 2 = somewhat agree, 3 = absolutely agree on each item. Source: https://doi.org/10.1007/s10597-013-9619-1

```{r pulling bhs values, include=FALSE}

## Pulling mfq variables

# vars_bhs <- data_mfq_scored %>%
#   select(contains("bhs")) %>%
#   head() %>%
#   print()

```

Mutating bhs character values to numeric values

```{r mutating bhs values}

## Mutating character values to numeric across multiple bhs columns 

data_bhs_mutated <- data_mfq_scored %>%
  mutate(across(contains("bhs_") & -contains("tim"), 
                ~ case_when(
                    . == "Absolutely disagree" ~ 0,
                    . == "Somewhat disagree" ~ 1,
                    . == "Somewhat agree" ~ 2,
                    . == "Absolutely agree" ~ 3,
                    TRUE ~ NA_real_ # Handle other cases with NA
                )))

```

