---
title: "yes_data_cleaning"
author: "Mallory Dobias"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    toc: yes
    toc_depth: 5
    toc_float:
      collapsed: no
      smooth_scroll: no
  pdf_document:
    toc: yes
    toc_depth: '5'
  word_document:
    toc: yes
    toc_depth: '5'
geometry: margin=0.50in
---

```{r setup, include=FALSE}

## Setting up R

knitr::opts_chunk$set(echo = TRUE)

if(!require(pointblank)){install.packages('pointblank')}
library(pointblank) # installing and calling pointblank package early in Markdown

validate_rmd(summary = TRUE, log_to_file = FALSE) # allowing validation from pointblank package within Markdown

```

```{r, include=FALSE}

if(!require(here)){install.packages('here')}
library(here)

if(!require(janitor)){install.packages('janitor')}
library(janitor)

if(!require(dplyr)){install.packages('dplyr')}
library(dplyr)

if(!require(stringr)){install.packages('stringr')}
library(stringr)

if(!require(pointblank)){install.packages('pointblank')}
library(pointblank)

if(!require(tidyr)){install.packages('tidyr')}
library(tidyr)

if(!require(readr)){install.packages('readr')}
library(readr)

```

# Data Reading

```{r reading in data}

## Reading in NU English survey data

library(here) # places working directory at root of project folder

raw_data <- read.csv(here("data", "raw", "Project YES - ENGLISH AND SPANISH - Activity Only_January 2, 2025_12.03.csv")) # directing to data from project folder root

# head(raw_data)

```

# Data Cleaning

## Cleaning Var Names

Removing Qualtrics metadata rows

```{r removing first two Qualtrics metadata rows}

## Removing first two rows with Qualtrics metadata

data_sliced <- raw_data %>%
  dplyr::slice(-1,-2)

# head(data_sliced) # looks good

```

```{r checking metadata removal setup, include = FALSE}

## checking Qualtrics metadata rows are removed

# pulling row 3 of raw data to compare to row 1 of sliced data

test_sliced_pre <- raw_data %>%
  dplyr::slice(3) # pulling pre-removal row 3 for comparison

test_sliced_post <- data_sliced %>%
  dplyr::slice(1) # pulling post-removal row 1 for comparison

# combining into one dataframe

test_sliced <- rbind(test_sliced_pre, test_sliced_post)

```

Checking metadata rows removed correctly

```{r checking metadata removal, validate = TRUE}

test_sliced %>%
  pointblank::col_vals_equal(vars(ResponseId), 
                             vars(ResponseId)) # passes if third row response is now first row

```

Converting to tidy var names

```{r creating tidy names}

## Cleaning var names

data_named <- data_sliced %>%
  janitor::clean_names()

# head(data_named)

```

Checking var names are tidy

```{r checking tidy names setup}

## checking var names are now tidy

test_var_names <- tibble(col_names = names(data_named)) # creates data frame of names

test_var_names %>%
  filter(stringr::str_detect(col_names, "[A-Z]")) %>%
  nrow() == 0 # true if zero col names contain capital letters

```

Renaming "pi_" (post-intervention) variable prefixes to "post_" prefixes, for clarity

```{r renaming pi vars with post label}

## Renaming "pi_" (post-intervention) vars with "post_" prefix for clarity

data_post_named <- data_named %>%
  rename_with(~ str_replace(., "^pi_", "post_"))

```

```{r checking post var naming setup, include = FALSE}

## checking vars are renamed correctly

# extracting characters following "pre_" in column names before renaming

test_pre_rename <- tibble(var_names = names(data_named)) %>%
  filter(str_detect(var_names, "^pre_") & !str_detect(var_names, "tim")) %>% # keep non-time rows with "pre_" prefix
  mutate(pre_renamed = str_extract(var_names, "(?<=pre_).*")) %>% # extract characters following "pre_"
  select(-var_names)

# extracting characters following "post_" in column names after renaming

test_post_rename <- tibble(var_names = names(data_post_named)) %>%
  filter(str_detect(var_names, "^post_") & !str_detect(var_names, "tim")) %>% # keep non-time rows with "post_" prefix
  mutate(post_renamed = str_extract(var_names, "(?<=post_).*")) %>% # extract characters following "post_"
  select(-var_names)

# combining to one data frame for comparison

test_rename <- cbind(test_pre_rename, test_post_rename)

```

Checking columns renamed correctly

```{r checking post var renaming, validate = TRUE}

test_rename %>%
  pointblank::col_vals_equal(vars(pre_renamed), 
                             vars(post_renamed)) # passes if renamed columns are identical to original cols

```

Renaming "prog_" (program) variable prefixes to "ssi_" prefixes, for clarity

```{r renaming prog vars with ssi label}

## Renaming "prog_" (program) variable prefixes to "ssi_" prefixes, for clarity

data_prog_named <- data_post_named %>%
  rename_with(~ str_replace(., "^prog_", "ssi_"))

```

```{r checking prog var naming setup, include = FALSE}

## checking vars are renamed correctly

# extracting characters following "prog_" in column names before renaming

test_pre_rename <- tibble(var_names = names(data_post_named)) %>%
  filter(str_detect(var_names, "^prog_") & !str_detect(var_names, "tim")) %>% # keep non-time rows with "prog_" prefix
  mutate(pre_renamed = str_extract(var_names, "(?<=prog_).*")) %>% # extract characters following "prog_"
  select(-var_names)

# extracting characters following "ssi_" in column names after renaming

test_post_rename <- tibble(var_names = names(data_prog_named)) %>%
  filter(str_detect(var_names, "^ssi_") & !str_detect(var_names, "tim") & !str_detect(var_names, "choose")
         & !str_detect(var_names, "source")) %>% # keep non-time rows with "ssi_" prefix
  mutate(post_renamed = str_extract(var_names, "(?<=ssi_).*")) %>% # extract characters following "ssi_"
  select(-var_names)

# combining to one data frame for comparison

test_rename <- cbind(test_pre_rename, test_post_rename)

```

Checking columns renamed correctly

```{r checking prog var renaming, validate = TRUE}

test_rename %>%
  pointblank::col_vals_equal(vars(pre_renamed), 
                             vars(post_renamed)) # passes if renamed columns are identical to original cols

```

Renaming "into_" (typo) variable prefixes to "intro_" prefixes, for clarity and consistency

```{r renaming intro vars}

## Renaming "prog_" (program) variable prefixes to "ssi_" prefixes, for clarity

data_intro_named <- data_prog_named %>%
  rename_with(~ str_replace(., "^into_", "intro_"))

```

```{r checking intro var naming setup, include = FALSE}

## checking vars are renamed correctly

# extracting characters following "into_" in column names before renaming

test_pre_rename <- tibble(var_names = names(data_prog_named)) %>%
  filter(str_detect(var_names, "^into_")) %>% # keep rows with "into_" prefix
  mutate(pre_renamed = str_extract(var_names, "(?<=into_).*")) %>% # extract characters following "into_"
  select(-var_names)

# extracting characters following "intro_" in column names after renaming

test_post_rename <- tibble(var_names = names(data_intro_named)) %>%
  filter(str_detect(var_names, "^intro_")) %>% # keep rows with "intro_" prefix
  mutate(post_renamed = str_extract(var_names, "(?<=intro_).*")) %>% # extract characters following "intro_"
  select(-var_names)

# combining to one data frame for comparison

test_rename <- cbind(test_pre_rename, test_post_rename)

```

Checking columns renamed correctly

```{r checking intro var renaming, validate = TRUE}

test_rename %>%
  pointblank::col_vals_equal(vars(pre_renamed), 
                             vars(post_renamed)) # passes if renamed columns are identical to original cols

```

```{r removing ao string from var names}

## Removing '_ao' from all column names, regardless of where it appears

data_ao_removed <- data_intro_named %>%
  rename_with(~ str_replace_all(., "_ao", ""))

```

Checking columns renamed correctly

```{r checking ao string removed, validate = TRUE}

## Checking if the _ao string was removed from all vars correctly

validation_result <- 
  data_ao_removed %>%
  select(contains("_ao")) %>%
  expect_col_count_match(0)

## Printing the validation result - runs if passes

validation_result

```

Renaming individual items for clarity

```{r renaming individual vars}

## Renaming individual vars for clarity

data_items_named <- data_ao_removed %>%
  rename(perc_progress = progress,
         ssi_started = project_started,
         ssi_completed = project_completed,
         ssi_choice = ssi_which_activi,
         abc_q_6_text = abc_q_6, # adding _text to qual entry vars in ssis
         abc_q_8_text = abc_q_8,
         abc_q_37_text = abc_slide_37,
         proj_care_q_1_text = proj_care_q_1,
         proj_care_q_2_text = proj_care_q_2,
         proj_care_q_4_text = proj_care_q_4,
         proj_care_q_5_text = proj_care_q_5,
         proj_care_q_6_text = proj_care_q_6,
         proj_pers_q_1_text = proj_pers_q_1,
         proj_pers_q_2_text = proj_pers_q_2,
         proj_pers_q_3_text = proj_pers_q_3,
         abc_q_1_1_num = abc_q_1_1, # adding _num to numeric vars in ssis
         abc_q_3_1_num = abc_q_3_1)

```

## Cleaning Values

Replacing empty cells with NAs

```{r replacing empty cells with nas}

## Replacing empty cells with NAs

data_w_nas <- data_items_named %>%
  mutate_all(~ na_if(., ""))

```

```{r checking nas setup, include=FALSE}

## Creating data frames of ids of folks with missing data for ssi_started pre and post nas

test_pre_nas <- data_items_named %>%
  filter(ssi_started == "") %>%
  select(response_id) # ids with missing data pre nas ("")

test_post_nas <- data_w_nas %>%
  filter(is.na(ssi_started)) %>%
  select(response_id) # ids with missing data post nas (NA)

# Get the unique names (names not in both data frames)

test_nas <- bind_rows(test_pre_nas, test_post_nas) %>%
  group_by(response_id) %>%
  filter(n() == 1)
         
```

Checking that nas mutation happened correctly

```{r checking reorg did not lose cols}

## Checking ids with empty data for age are identical to ids with nas for age

n_ids_missing_age <- test_nas %>%
  nrow() == 0

n_ids_missing_age # true if ids missing age data are identical pre and post age mutation

```

Cleaning values in start_date and end_date columns

```{r cleaning date columns}

data_dated <- data_w_nas %>%
  mutate(start_date = str_remove(start_date, " .*$"),
         end_date = str_remove(end_date, " .*$"))

```

```{r checking date cleaning setup, include=FALSE}

## Checking dates are identical pre to post cleaning

# pulling first 10 characters in date column pre-cleaning

test_pre_date <- data_w_nas %>%
  mutate(start_date = str_sub(start_date, 1, 10),
    end_date = str_sub(end_date, 1, 10)) %>%
  select(pre_start_date = start_date, pre_end_date = end_date)

# pulling first 10 characters in date column pre-cleaning

test_post_date <- data_dated %>%
  mutate(start_date = str_sub(start_date, 1, 10),
    end_date = str_sub(end_date, 1, 10)) %>%
  select(post_start_date = start_date, post_end_date = end_date)

# binding cols together for comparison

test_dates <- cbind(test_pre_date, test_post_date)

```

Checking dates cleaned correctly

```{r checking date cleaning, validate = TRUE}

## checking start_date and end_date are identical pre to post cleaning

test_dates %>%
  pointblank::col_vals_equal(vars(pre_start_date), 
                             vars(post_start_date)) %>%
  pointblank::col_vals_equal(vars(pre_end_date), 
                             vars(post_end_date)) # passes if dates are identical pre to post cleaning

```

Combining the ssi_choice and ssi_choose vars

```{r combining ssi_choice and ssi_chose}

## Cleaning ssi_choice variable

data_ssi_choice_cleaned <- data_dated %>%
  mutate(ssi_choice = case_when(str_detect(ssi_choice, "A.B.C") ~ "ABC Project",
                                str_detect(ssi_choice, "C.A.R.E") ~ "Project CARE",
                                str_detect(ssi_choice, "Personality") ~ "Project Personality",
                                TRUE ~ ssi_choice))

```

```{r checking ssi_choice setup, include=FALSE}

## Creating data frames of ids of folks who chose an ssi

test_pre_mutate_count <- data_dated %>%
  filter(str_detect(ssi_choice, "A.B.C")) %>%
  select(pre_response_id = response_id) # ids of non na responses to choice vars pre mutation

test_post_mutate_count <- data_ssi_choice_cleaned %>%
  filter(str_detect(ssi_choice, "ABC Project")) %>%
  select(pre_response_id = response_id) # ids of non na responses to choice vars pre mutation

# combining data frames for comparison

test_mutate_count <- cbind(test_pre_mutate_count, test_post_mutate_count)

```

Checking ssi_choice variables cleaned correctly

```{r checking ssi_choice cleaning, validate = TRUE}

## checking counts align pre to post cleaning

test_mutate_count %>%
  pointblank::col_vals_equal(vars(pre_response_id), 
                             vars(post_response_id)) # passes if ids are identical pre to post cleaning

```

## Subsetting to English

```{r subsetting to english responses and vars}

data_english <- data_ssi_choice_cleaned %>%
  filter(language == "english")

```

```{r checking language subset, include=FALSE}

## Creating data frames of ids of folks with English embedded data and responses to language item

test_embedded_count <- data_ssi_choice_cleaned %>%
  filter(language == "english") %>%
  select(pre_response_id = response_id) # ids of embedded data set to english

test_item_count <- data_ssi_choice_cleaned %>%
  filter(select_language == "English (Inglés)") %>%
  select(pre_response_id = response_id) # ids of responses who selected english

# combining data frames for comparison

test_mutate_count <- cbind(test_embedded_count, test_item_count)

```

Checking language subset correctly

```{r checking language filter, validate = TRUE}

## checking counts align pre to post cleaning

test_mutate_count %>%
  pointblank::col_vals_equal(vars(pre_response_id), 
                             vars(post_response_id)) # passes if ids are identical pre to post cleaning

```

```{r selecting english vars}

## Selecting only english vars

data_english_vars_only <- data_english %>%
  select(-starts_with("s_"), -ends_with("_s"))

```

## Calculating Composites

### PFS

Calculating program feedback scores using the Program feedback scale (measured post SSI). Row mean scores range from 1 to 5, where 1 = really disagree, and 5 = really agree. Source: https://mental.jmir.org/2020/6/e20513

Mutating pfs character values to numeric values

```{r mutating pfs values}

## Mutating character values to numeric across multiple pfs columns 

data_pfs_mutated <- data_english_vars_only %>%
  mutate(across(contains("pfs_") & -contains("like") & 
               -contains("change") &
               -contains("more") &
               -contains("tim"),
                ~ case_when(
                    . == "Really Disagree" ~ 1,
                    . == "Disagree" ~ 2,
                    . == "Neutral" ~ 3,
                    . == "Agree" ~ 4,
                    . == "Really Agree" ~ 5,
                    TRUE ~ NA_real_ # Handle other cases with NA
                )))
 
```

Checking pfs variables are now numeric class

```{r pfs numeric class check}

## Checking that all pfs vars are numeric class

data_pfs_mutated %>%
  select(contains("pfs") & 
         -contains("like") & 
         -contains("change") &
         -contains("more") &
         -contains("tim"),) %>%
  summarise(across(everything(), ~ class(.x)[1])) %>%
  tidyr::pivot_longer(cols = everything(), names_to = "variable", values_to = "class")


```

```{r writing a function to setup future mutation checks, include=FALSE}

# Custom function to compare value counts before and after mutation
prep_mutation_check <- function(data_pre, data_post, var) {
  
  # Step 1: Create pre-mutation counts
  test_pre <- data_pre %>%
    select(var) %>%
    group_by(across(all_of(var))) %>%  # Group by the specified variable
    summarize(count = n(), .groups = "drop") %>%  # Count occurrences
    arrange(desc(count)) %>%  # Arrange in descending order
    rename_with(~ paste0(., "_pre"), everything())  # Rename columns for pre-mutation
  
  # Step 2: Create post-mutation counts
  test_post <- data_post %>%
    select(var) %>%
    mutate(across(all_of(var), ~ factor(., levels = unique(.)))) %>%  # Convert to factor
    group_by(across(all_of(var))) %>%  # Group by the specified variable
    summarize(count = n(), .groups = "drop") %>%  # Count occurrences
    arrange(desc(count)) %>%  # Arrange in descending order
    rename_with(~ paste0(., "_post"), everything())  # Rename columns for post-mutation
  
  # Step 3: Combine pre and post counts into one dataframe for comparison
  test_comparison <- bind_cols(test_pre, test_post)
  
  return(test_comparison)
}

```

```{r using function to setup pfs mutation check, include=FALSE}

# `data_hate_scored` is the pre-mutation dataframe
# and `data_pfs_mutated` is the post-mutation dataframe,
# and we want to compare the value counts of "pfs_1":

test_mutate <- prep_mutation_check(data_english_vars_only, data_pfs_mutated, "pfs_1")

```

Checking pfs response counts are identical from pre to post mutation

```{r pfs response counts validation, validate = TRUE}

test_mutate %>%
  pointblank::col_vals_equal(vars(count_pre), 
                             vars(count_post)) # passes if response counts are identical pre to post mutation

```

Calculating pfs row means

```{r calculating pfs scores}

## Calculating composite scores

data_pfs_scored <- data_pfs_mutated %>%
  # Calculate the row means of relevant pfs_ columns, excluding specific patterns
  mutate(pfs_mean = rowMeans(across(contains("pfs_") & 
                                    !contains("like") & 
                                    !contains("change") &
                                    !contains("more") &
                                    !contains("tim")), na.rm = TRUE)) %>%
  # Replace NaN values with NA_real_ across relevant pfs columns
  mutate(across(contains("pfs_") &
                !contains("like") & 
                !contains("change") &
                !contains("more") &
                !contains("tim"),
                ~ case_when(is.nan(.x) ~ NA_real_,  # Replace NaN with NA
                            TRUE ~ .x))) # Keep other values unchanged

```

```{r checking pfs score setup, include = FALSE}

## checking composite calculation was correct

# manually calculating means to compare methods

test_pfs_calc <- data_pfs_scored %>%
  select(response_id, starts_with("pfs") & 
                                    !contains("like") & 
                                    !contains("change") &
                                    !contains("more") &
                                    !contains("tim")) %>%
  mutate(manual_mean = rowMeans(select(., pfs_1, pfs_2, pfs_3, pfs_4, pfs_5, pfs_6, pfs_7), na.rm = TRUE)) %>%
  mutate(across(contains("manual"), ~ case_when(is.nan(.) ~ NA_real_, # Replace NaN with NA
                                        TRUE ~ .))) %>% # Keep other values unchanged
  select(pfs_mean, manual_mean)

```

Checking pfs mean is identical across averaging methods

```{r pfs scoring validation, validate = TRUE}

test_pfs_calc %>%
  pointblank::col_vals_equal(vars(pfs_mean), 
                             vars(manual_mean), 
                             na_pass = TRUE) # passes if means are identical across averaging methods

```

## Reorganizing Columns

Dropping unnecessary columns

```{r dropping cols}

data_col_dropped <- data_pfs_scored %>%
  select(-status, -contains("duration"), -contains("recipient"), -contains("external"), -contains("distribution"), 
         -recorded_date, -user_language) %>% # dropping unnecessary qualtrics metadata
  select(-select_language) # dropping redundant language column

```

Adding cols to document language and data source

```{r adding cols}

data_col_added <- data_col_dropped %>%
  mutate(language = "english",
         data_id = "d03")

```

Reorganizing column order to place important columns first, and then to follow chronological order

```{r reorganizing col order}

## Reorganizing cols to prioritize important cols first and then chronological order

data_reorged <- data_col_added %>%
  select(response_id, language, data_id, contains("date"), ip_address, starts_with("location"),
         perc_progress, finished, ssi_choice,
         starts_with("volunteer_") & !contains("tim"),
         contains("reliant"), contains("san_antonio"),
         ssi_started,
         starts_with("abc") & !contains("tim"),
         starts_with("proj_care") & !contains("tim"),
         starts_with("proj_pers") & !contains("tim"),
         starts_with("proj_rise") & !contains("tim"),
         ssi_completed,
         starts_with("advice") & !contains("tim"),
         starts_with("pfs_") & !contains("tim"),
         starts_with("perc_change_") & !contains("tim"),
         contains("tim") & starts_with("select_lang"),
         contains("tim") & starts_with("intro"),
         contains("tim") & starts_with("into"),
         contains("tim") & starts_with("volunteer"),
         contains("tim") & starts_with("dem"),
         contains("tim") & starts_with("ssi"),
         contains("tim") & starts_with("mfq"),
         contains("tim") & starts_with("pre_"),
         contains("tim") & starts_with("abc"),
         contains("tim") & starts_with("proj_care"),
         contains("tim") & starts_with("proj_rise"),
         contains("tim") & starts_with("proj_pers"),
         contains("tim") & starts_with("advice_"),
         contains("tim") & starts_with("pfs_"),
         contains("tim") & starts_with("perc_change_"),
         contains("tim") & starts_with("post_"))

```

```{r checking col order setup, include=FALSE}

## creating data frames of names pre and post reorg for comparison

pre_reorg_names <- data_col_added %>%
  names() # names pre

post_reorg_names <- data_reorged %>%
  names() # names post

# Create tibbles for easier manipulation

pre_reorg_df <- tibble(names = pre_reorg_names)
post_reorg_df <- tibble(names = post_reorg_names)

# Get the unique names (names not in both data frames)

test_reorg <- bind_rows(pre_reorg_df, post_reorg_df) %>%
  group_by(names) %>%
  filter(n() == 1)
         
```

Checking that reorganizing happened correctly (i.e., no columns dropped) 

```{r checking reorg did not lose cols 2}

## Checking if df with any columns unique to pre-reorg data is empty (i.e., no columns dropped)

n_cols_dropped <- test_reorg %>%
  nrow() == 0

n_cols_dropped # true if no columns were dropped by reorg

```

## Converting Classes

Converting variable classes

```{r mutating var classes}

data_classes_mutated <- data_reorged %>%
  select(-ip_address, -starts_with("location")) %>%
  mutate(language = as.factor(language),
         data_id = as.factor(data_id),
         perc_progress = as.numeric(perc_progress),
         finished = as.factor(finished),
         ssi_choice = as.factor(ssi_choice),
         ssi_started = as.factor(ssi_started),
         ssi_completed = as.factor(ssi_completed),
         advice_permission = as.factor(advice_permission)) %>%
  mutate(across(contains("date"), as.Date)) %>%
  mutate(across(contains("_q_") & ends_with("_num"), as.numeric)) %>% # ssi qual questions
  mutate(across(contains("_q_") & -ends_with("_text") & -ends_with("_num") &
                -ends_with("_rank") & -ends_with("_group"), as.factor)) %>% # ssi qual questions
  mutate(across(contains("commit"), as.factor)) %>%
  mutate(across(contains("perc_change"), as.factor)) %>%
  mutate(across(contains("_tim_"), as.numeric))

```

# Saving Data

```{r saving cleaned data}

data <- data_classes_mutated

# Define the folder and file path

folder_path <- here("data", "clean")  # Path to my desired folder
file_name <- "yes_data_d03_english_clean.Rds"
full_path <- file.path(folder_path, file_name)

# Save the dataframe to the .csv file

saveRDS(data, full_path)

```


