---
title: "yes_data_cleaning"
author: "Mallory Dobias"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    toc: yes
    toc_depth: 5
    toc_float:
      collapsed: no
      smooth_scroll: no
  pdf_document:
    toc: yes
    toc_depth: '5'
  word_document:
    toc: yes
    toc_depth: '5'
geometry: margin=0.50in
---

```{r setup, include=FALSE}

## Setting up R

knitr::opts_chunk$set(echo = TRUE)

if(!require(pointblank)){install.packages('pointblank')}
library(pointblank) # installing and calling pointblank package early in Markdown

validate_rmd(summary = TRUE, log_to_file = FALSE) # allowing validation from pointblank package within Markdown

```

```{r, include=FALSE}

if(!require(here)){install.packages('here')}
library(here)

if(!require(janitor)){install.packages('janitor')}
library(janitor)

if(!require(dplyr)){install.packages('dplyr')}
library(dplyr)

if(!require(stringr)){install.packages('stringr')}
library(stringr)

if(!require(pointblank)){install.packages('pointblank')}
library(pointblank)

if(!require(tidyr)){install.packages('tidyr')}
library(tidyr)

if(!require(readr)){install.packages('readr')}
library(readr)

if(!require(datadictionary)){install.packages('datadictionary')}
library(datadictionary)

if(!require(codebook)){install.packages('codebook')}
library(codebook)

```

# Data Reading

```{r reading in data}

## Reading in NU Turkish survey data

library(here) # places working directory at root of project folder

raw_data <- read.csv(here("data", "raw", "NU - YES - Turkish - Questionnaires Optional_March 13, 2025_11.27.csv")) # directing to data from project folder root

# head(raw_data)

```

# Data Cleaning

## Cleaning Var Names

Removing Qualtrics metadata rows

```{r removing first two Qualtrics metadata rows}

## Removing first two rows with Qualtrics metadata

data_sliced <- raw_data %>%
  dplyr::slice(-1,-2)

# head(data_sliced) # looks good

```

```{r checking metadata removal setup, include = FALSE}

## checking Qualtrics metadata rows are removed

# pulling row 3 of raw data to compare to row 1 of sliced data

test_sliced_pre <- raw_data %>%
  dplyr::slice(3) # pulling pre-removal row 3 for comparison

test_sliced_post <- data_sliced %>%
  dplyr::slice(1) # pulling post-removal row 1 for comparison

# combining into one dataframe

test_sliced <- rbind(test_sliced_pre, test_sliced_post)

```

Checking metadata rows removed correctly

```{r checking metadata removal, validate = TRUE}

test_sliced %>%
  pointblank::col_vals_equal(vars(ResponseId), 
                             vars(ResponseId)) # passes if third row response is now first row

```

Converting to tidy var names

```{r creating tidy names}

## Cleaning var names

data_named <- data_sliced %>%
  janitor::clean_names()

# head(data_named)

```

Checking var names are tidy

```{r checking tidy names setup}

## checking var names are now tidy

test_var_names <- tibble(col_names = names(data_named)) # creates data frame of names

test_var_names %>%
  filter(stringr::str_detect(col_names, "[A-Z]")) %>%
  nrow() == 0 # true if zero col names contain capital letters

```

Removing "t_" prefix, for later joins

```{r removing t prefix}

## Removing "t_" prefix, for later joins

data_t_removed <- data_named %>%
  rename_with(~ str_replace(., "^t_", ""))

```

```{r checking removal of t prefix setup, include = FALSE}

## checking vars are renamed correctly

test_pre_t_removal <- tibble(var_names = names(data_named)) %>%
  filter(str_detect(var_names, "^t_")) %>%
  nrow()

test_post_t_removal <- tibble(var_names = names(data_t_removed)) %>%
  filter(str_detect(var_names, "^t_")) %>%
  nrow()

```

Checking columns removed t prefix correctly

```{r checking t prefix removal, validate = TRUE}

test_post_t_removal == 0 # true if no cols start with "t_" prefix

```

Renaming "pi_" (post-intervention) variable prefixes to "post_" prefixes, for clarity

```{r renaming pi vars with post label}

## Renaming "pi_" (post-intervention) vars with "post_" prefix for clarity

data_post_named <- data_t_removed %>%
  rename_with(~ str_replace(., "^pi_", "post_"))

```

```{r checking post var naming setup, include = FALSE}

## checking vars are renamed correctly

# extracting characters following "pre_" in column names before renaming

test_pre_rename <- tibble(var_names = names(data_t_removed)) %>%
  filter(str_detect(var_names, "^pi_") & !str_detect(var_names, "tim")) %>% # keep non-time rows with "pi_" prefix
  mutate(pre_renamed = str_extract(var_names, "(?<=pi_).*")) %>% # extract characters following "pi_"
  select(-var_names)

# extracting characters following "post_" in column names after renaming

test_post_rename <- tibble(var_names = names(data_post_named)) %>%
  filter(str_detect(var_names, "^post_") & !str_detect(var_names, "tim")) %>% # keep non-time rows with "post_" prefix
  mutate(post_renamed = str_extract(var_names, "(?<=post_).*")) %>% # extract characters following "post_"
  select(-var_names)

# combining to one data frame for comparison

test_rename <- cbind(test_pre_rename, test_post_rename)

```

Checking columns renamed correctly

```{r checking post var renaming, validate = TRUE}

test_rename %>%
  pointblank::col_vals_equal(vars(pre_renamed), 
                             vars(post_renamed)) # passes if renamed columns are identical to original cols

```

Renaming "prog_" (program) variable prefixes to "ssi_" prefixes, for clarity

```{r renaming prog vars with ssi label}

## Renaming "prog_" (program) variable prefixes to "ssi_" prefixes, for clarity

data_prog_named <- data_post_named %>%
  rename_with(~ str_replace(., "^prog_", "ssi_"))

```

```{r checking prog var naming setup, include = FALSE}

## checking vars are renamed correctly

# extracting characters following "prog_" in column names before renaming

test_pre_rename <- tibble(var_names = names(data_post_named)) %>%
  filter(str_detect(var_names, "^prog_") & !str_detect(var_names, "tim")) %>% # keep non-time rows with "prog_" prefix
  mutate(pre_renamed = str_extract(var_names, "(?<=prog_).*")) %>% # extract characters following "prog_"
  select(-var_names)

# extracting characters following "ssi_" in column names after renaming

test_post_rename <- tibble(var_names = names(data_prog_named)) %>%
  filter(str_detect(var_names, "^ssi_") & !str_detect(var_names, "tim") & !str_detect(var_names, "choose")
         & !str_detect(var_names, "source")) %>% # keep non-time rows with "ssi_" prefix
  mutate(post_renamed = str_extract(var_names, "(?<=ssi_).*")) %>% # extract characters following "ssi_"
  select(-var_names)

# combining to one data frame for comparison

test_rename <- cbind(test_pre_rename, test_post_rename)

```

Checking columns renamed correctly

```{r checking prog var renaming, validate = TRUE}

test_rename %>%
  pointblank::col_vals_equal(vars(pre_renamed), 
                             vars(post_renamed)) # passes if renamed columns are identical to original cols

```

Renaming individual items for clarity

```{r renaming individual vars}

## Renaming individual vars for clarity

data_items_named <- data_prog_named %>%
  rename(dem_receive_support = dem_receive_suppor,
         pre_perceived_control = pre_control_1, 
         post_perceived_control = post_control_1,
         perc_progress = progress, 
         volunteer_questionnaires_yn = volunteer,
         ssi_started = project_started,
         ssi_completed = project_completed,
         ssi_choice = ssi_which_activit,
         ssi_prev_completed = ssi_prev_activity,
         ssi_prev_completed_name = ssi_prev_activi_2,
         ssi_times_completed_personality = ssi_proj_personal,
         ssi_times_completed_care = ssi_proj_care,
         ssi_times_completed_abc = ssi_proj_abc,
         abc_q_6_text = abc_q_6, # adding _text to qual entry vars in ssis
         abc_q_8_text = abc_q_8,
         abc_q_37_text = abc_slide_37,
         proj_care_q_1_text = proj_care_q_1,
         proj_care_q_2_text = proj_care_q_2,
         proj_care_q_4_text = proj_care_q_4,
         proj_care_q_5_text = proj_care_q_5,
         proj_care_q_6_text = proj_care_q_6,
         proj_pers_q_1_text = proj_pers_q_1,
         proj_pers_q_2_text = proj_pers_q_2,
         proj_pers_q_3_text = proj_pers_q_3,
         abc_q_1_1_num = abc_q_1_1, # adding _num to numeric vars in ssis
         abc_q_3_1_num = abc_q_3_1)

```

## Cleaning Values

Replacing empty cells with NAs

```{r replacing empty cells with nas}

## Replacing empty cells with NAs

data_w_nas <- data_items_named %>%
  mutate_all(~ na_if(., ""))

```

```{r checking nas setup, include=FALSE}

## Creating data frames of ids of folks with missing data for age pre and post nas

test_pre_nas <- data_items_named %>%
  filter(dem_age == "") %>%
  select(response_id) # ids with missing data pre nas ("")

test_post_nas <- data_w_nas %>%
  filter(is.na(dem_age)) %>%
  select(response_id) # ids with missing data post nas (NA)

# Get the unique names (names not in both data frames)

test_nas <- bind_rows(test_pre_nas, test_post_nas) %>%
  group_by(response_id) %>%
  filter(n() == 1)
         
```

Checking that nas mutation happened correctly

```{r checking reorg did not lose cols}

## Checking ids with empty data for age are identical to ids with nas for age

n_ids_missing_age <- test_nas %>%
  nrow() == 0

n_ids_missing_age # true if ids missing age data are identical pre and post age mutation

```

Cleaning values in start_date and end_date columns

```{r cleaning date columns}

data_dated <- data_w_nas %>%
  mutate(start_date = str_remove(start_date, " .*$"),
         end_date = str_remove(end_date, " .*$"))

```

```{r checking date cleaning setup, include=FALSE}

## Checking dates are identical pre to post cleaning

# pulling first 10 characters in date column pre-cleaning

test_pre_date <- data_w_nas %>%
  mutate(start_date = str_sub(start_date, 1, 10),
    end_date = str_sub(end_date, 1, 10)) %>%
  select(pre_start_date = start_date, pre_end_date = end_date)

# pulling first 10 characters in date column pre-cleaning

test_post_date <- data_dated %>%
  mutate(start_date = str_sub(start_date, 1, 10),
    end_date = str_sub(end_date, 1, 10)) %>%
  select(post_start_date = start_date, post_end_date = end_date)

# binding cols together for comparison

test_dates <- cbind(test_pre_date, test_post_date)

```

Checking dates cleaned correctly

```{r checking date cleaning, validate = TRUE}

## checking start_date and end_date are identical pre to post cleaning

test_dates %>%
  pointblank::col_vals_equal(vars(pre_start_date), 
                             vars(post_start_date)) %>%
  pointblank::col_vals_equal(vars(pre_end_date), 
                             vars(post_end_date)) # passes if dates are identical pre to post cleaning

```

## Translating Factors

Creating separate prev_completed variables for easier counting

```{r cleaning prev_completed values}

## Creating separate prev_completed variables for easier counting

data_prev_completed_cleaned <- data_dated %>%
  mutate(ssi_prev_completed_abc = case_when(str_detect(ssi_prev_completed_name, "Proje 3: Eylem Değişim Getirir") ~ "yes",
                                            TRUE ~ NA_character_),
         ssi_prev_completed_care = case_when(str_detect(ssi_prev_completed_name, "Proje 2: Kendini Sevmek Ne işe Yarar?") ~ "yes",
                                            TRUE ~ NA_character_),
         ssi_prev_completed_pp = case_when(str_detect(ssi_prev_completed_name, "Proje 1: Beynimiz ve Biz") ~ "yes",
                                            TRUE ~ NA_character_)) %>%
  select(-ssi_prev_completed_name)
  
```

```{r checking prev_completed var cleaning setup, include=FALSE}

## Creating data frames of ids of folks who previously completed ABC

test_pre_mutate_count <- data_dated %>%
  filter(str_detect(ssi_prev_completed_name, "Proje 3: Eylem Değişim Getirir")) %>%
  select(pre_response_id = response_id) # ids of ABC participants pre mutation

test_post_mutate_count <- data_prev_completed_cleaned %>%
  filter(ssi_prev_completed_abc == "yes") %>%
  select(post_response_id = response_id) # ids of ABC participants post mutation

# combining data frames for comparison

test_mutate_count <- cbind(test_pre_mutate_count, test_post_mutate_count)

```

Checking ssi_prev_completed variables cleaned correctly

```{r checking prev_completed cleaning, validate = TRUE}

## checking prev_completed counts align pre to post cleaning

test_mutate_count %>%
  pointblank::col_vals_equal(vars(pre_response_id), 
                             vars(post_response_id)) # passes if yeses are identical pre to post cleaning

```

Combining the ssi_choice var

```{r combining ssi_choice and ssi_chose}

## Cleaning ssi_choice variable

data_ssi_choice_cleaned <- data_prev_completed_cleaned %>%
  mutate(ssi_choice = case_when(str_detect(ssi_choice, "Proje 3: Eylem Değişim Getirir") ~ "ABC Project",
                                str_detect(ssi_choice, "Proje 2: Kendini Sevmek Ne işe Yarar?") ~ "Project CARE",
                                str_detect(ssi_choice, "Proje 1: Beynimiz ve Biz") ~ "Project Personality",
                                TRUE ~ ssi_choice))

```

```{r checking ssi_choice setup, include=FALSE}

## Creating data frames of ids of folks who chose an ssi

test_pre_mutate_count <- data_prev_completed_cleaned %>%
  filter(!is.na(ssi_choice)) %>%
  select(pre_response_id = response_id) # ids of non na responses to choice vars pre mutation

test_post_mutate_count <- data_ssi_choice_cleaned %>%
  filter(!is.na(ssi_choice)) %>%
  select(pre_response_id = response_id) # ids of non na responses to choice vars pre mutation

# combining data frames for comparison

test_mutate_count <- cbind(test_pre_mutate_count, test_post_mutate_count)

```

Checking ssi_choice variables cleaned correctly

```{r checking ssi_choice cleaning, validate = TRUE}

## checking counts align pre to post cleaning

test_mutate_count %>%
  pointblank::col_vals_equal(vars(pre_response_id), 
                             vars(post_response_id)) # passes if ids are identical pre to post cleaning

```

Cleaning volunteer_questionnaires_yn

```{r mutating volunteer qs var for future joins, include=FALSE}

data_questionnaires_yn_cleaned <- data_ssi_choice_cleaned %>%
  mutate(volunteer_questionnaires_yn = case_when(volunteer_questionnaires_yn == "Evet" ~ "Yes",
                                                 volunteer_questionnaires_yn == "Hayır" ~ "No",
                                                 TRUE ~ volunteer_questionnaires_yn))

```

```{r writing a function to setup future mutation checks, include=FALSE}

# Custom function to compare value counts before and after mutation
prep_mutation_check <- function(data_pre, data_post, var) {
  
  # Step 1: Create pre-mutation counts
  test_pre <- data_pre %>%
    select(var) %>%
    group_by(across(all_of(var))) %>%  # Group by the specified variable
    summarize(count = n(), .groups = "drop") %>%  # Count occurrences
    arrange(desc(count)) %>%  # Arrange in descending order
    rename_with(~ paste0(., "_pre"), everything())  # Rename columns for pre-mutation
  
  # Step 2: Create post-mutation counts
  test_post <- data_post %>%
    select(var) %>%
    mutate(across(all_of(var), ~ factor(., levels = unique(.)))) %>%  # Convert to factor
    group_by(across(all_of(var))) %>%  # Group by the specified variable
    summarize(count = n(), .groups = "drop") %>%  # Count occurrences
    arrange(desc(count)) %>%  # Arrange in descending order
    rename_with(~ paste0(., "_post"), everything())  # Rename columns for post-mutation
  
  # Step 3: Combine pre and post counts into one dataframe for comparison
  test_comparison <- bind_cols(test_pre, test_post)
  
  return(test_comparison)
}

```

```{r testing function to setup future mutation checks, include=FALSE}

# Example usage:
# Assuming `data_ssi_choice_cleaned` is the pre-mutation dataframe
# and `data_questionnaires_yn_cleaned` is the post-mutation dataframe,
# and you want to compare the value counts of "volunteer_questionnaires_yn":

test_mutate <- prep_mutation_check(data_ssi_choice_cleaned, data_questionnaires_yn_cleaned, "volunteer_questionnaires_yn") # same output as test_mfq_mutate above

## will use the prep_mutation_check() function to set up mutation checks moving forward

```

Checking response counts are identical from pre to post mutation

```{r volunteer qs response counts validation, validate = TRUE}

test_mutate %>%
  pointblank::col_vals_equal(vars(count_pre), 
                             vars(count_post)) # passes if response counts are identical pre to post mutation

```

Cleaning dem_receive_support

```{r mutating support var for future joins, include=FALSE}

data_support_cleaned <- data_questionnaires_yn_cleaned %>%
  mutate(dem_receive_support = str_replace(dem_receive_support, "Evet, şu anda destek alıyorum", "yes, I am getting support now")) %>%
  mutate(dem_receive_support = str_replace(dem_receive_support, "Evet, geçmişte destek aldım", "yes, I have gotten support in the past")) %>%
  mutate(dem_receive_support = str_replace(dem_receive_support, "Hayır, ama geçmişte destek almak istedim", "no, but I have wanted to get support before")) %>%
  mutate(dem_receive_support = str_replace(dem_receive_support, "Hayır ve geçmişte destek almak istemedim", "no, and I have not wanted to get support before"))

```

```{r testing mutation support var, include=FALSE}

# Checking mutation

test_mutate <- prep_mutation_check(data_questionnaires_yn_cleaned, data_support_cleaned, "dem_receive_support")

```

Checking response counts are identical from pre to post mutation

```{r support response counts validation, validate = TRUE}

test_mutate %>%
  pointblank::col_vals_equal(vars(count_pre), 
                             vars(count_post)) # passes if response counts are identical pre to post mutation

```

Cleaning dem_age

```{r mutating age var for future joins, include=FALSE}

data_age_cleaned <- data_support_cleaned %>%
  mutate(dem_age = case_when(dem_age == "10 veya altı" ~ "10 or younger",
                             dem_age == "18 veya üzeri" ~ "18 or older",
                             TRUE ~ dem_age))

```

```{r testing mutation age var, include=FALSE}

# Checking mutation

test_mutate <- prep_mutation_check(data_support_cleaned, data_age_cleaned, "dem_age")

```

Checking response counts are identical from pre to post mutation

```{r age response counts validation, validate = TRUE}

test_mutate %>%
  pointblank::col_vals_equal(vars(count_pre), 
                             vars(count_post)) # passes if response counts are identical pre to post mutation

```

Cleaning dem_sex

```{r mutating sex var for future joins, include=FALSE}

data_sex_cleaned <- data_age_cleaned %>%
  mutate(dem_sex = case_when(dem_sex == "Erkek" ~ "Male",
                             dem_sex == "Kadın" ~ "Female",
                             dem_sex == "Diğer" ~ "Other",
                             TRUE ~ dem_sex))

```

```{r testing mutation sex var, include=FALSE}

# Checking mutation

test_mutate <- prep_mutation_check(data_age_cleaned, data_sex_cleaned, "dem_sex")

```

Checking response counts are identical from pre to post mutation

```{r sex counts validation, validate = TRUE}

test_mutate %>%
  pointblank::col_vals_equal(vars(count_pre), 
                             vars(count_post)) # passes if response counts are identical pre to post mutation

```

Cleaning ssi_prev_completed

```{r mutating prev completed var for future joins, include=FALSE}

data_prev_completed_translated <- data_sex_cleaned %>%
  mutate(ssi_prev_completed = case_when(ssi_prev_completed == "Evet" ~ "Yes",
                                        ssi_prev_completed == "Hayır" ~ "No",
                                        TRUE ~ ssi_prev_completed))

```

```{r testing mutation prev completed var, include=FALSE}

# Checking mutation

test_mutate <- prep_mutation_check(data_sex_cleaned, data_prev_completed_translated, "ssi_prev_completed")

```

Checking response counts are identical from pre to post mutation

```{r prev completed counts validation, validate = TRUE}

test_mutate %>%
  pointblank::col_vals_equal(vars(count_pre), 
                             vars(count_post)) # passes if response counts are identical pre to post mutation

```

Cleaning commit vars

```{r mutating commit vars for future joins, include=FALSE}

data_commits_cleaned <- data_prev_completed_translated %>%
  mutate(across(contains("commit"), 
                ~ case_when(. == "Evet" ~ "Yes",
                            . == "Hayır" ~ "No",
                            TRUE ~ .)))

```

```{r testing mutation commit vars, include=FALSE}

# Checking mutation

test_mutate_abc <- prep_mutation_check(data_prev_completed_translated, data_commits_cleaned, "abc_commit")

test_mutate_care <- prep_mutation_check(data_prev_completed_translated, data_commits_cleaned, "proj_care_commit")

test_mutate_pp <- prep_mutation_check(data_prev_completed_translated, data_commits_cleaned, "proj_pers_commit")

```

Checking response counts are identical from pre to post mutation

```{r commit vars counts validation, validate = TRUE}

test_mutate_abc %>%
  pointblank::col_vals_equal(vars(count_pre), 
                             vars(count_post)) # passes if response counts are identical pre to post mutation

test_mutate_care %>%
  pointblank::col_vals_equal(vars(count_pre), 
                             vars(count_post)) # passes if response counts are identical pre to post mutation

test_mutate_pp %>%
  pointblank::col_vals_equal(vars(count_pre), 
                             vars(count_post)) # passes if response counts are identical pre to post mutation

```

## Calculating Composites

### Depression

Calculating depressive symptoms using the Short Form Mood and Feelings Questionnaire (only measured pre-SSI). Scores range from 0 to 26, where 0 = not true, 1 = sometimes, 2 = true. Scores of 12 or higher suggest depression. Source: https://doi.org/10.1002/mpr.1610

Mutating mfq character values to numeric values

```{r mutating mfq values}

## Mutating character values to numeric across multiple mfq columns 

data_mfq_mutated <- data_ssi_choice_cleaned %>%
  mutate(across(starts_with("mfq_") & -contains("tim"), 
                ~ case_when(
                    . == "Doğru değil" ~ 0,
                    . == "Bazen" ~ 1,
                    . == "Doğru" ~ 2,
                    TRUE ~ NA_real_ # Handle other cases with NA
                )))

```

Checking mfq variables are now numeric class

```{r mfq numeric class check}

## Checking that all mfq vars are numeric class

data_mfq_mutated %>%
  select(contains("mfq") & -contains("tim")) %>%
  summarise(across(everything(), ~ class(.x)[1])) %>%
  tidyr::pivot_longer(cols = everything(), names_to = "variable", values_to = "class")


```

```{r checking mfq mutation setup, include=FALSE}

## checking mutation happened correctly - setting up for pointblank tests

# creating df of count for each response, pre mutation

test_mfq_pre <- data_post_named %>%
  group_by(mfq_1) %>% # grouping by response
  summarize(count = n()) %>% # counting responses pre-mutation
  arrange(desc(count)) %>% # arranged in descending count order
  rename(mfq_1_pre = mfq_1, count_pre = count) # renaming before combining to one df

# creating df of count for each response, post mutation

test_mfq_post <- data_mfq_mutated %>%
  mutate(across(starts_with("mfq_"), 
                ~ factor(., levels = unique(.)))) %>% # converting to factor class so we can count
  group_by(mfq_1) %>% # grouping by response
  summarize(count = n()) %>% # counting responses post-mutation
  arrange(desc(count)) %>% # arranged in descending count order
  rename(mfq_1_post = mfq_1, count_post = count) # renaming before combining to one df

# creating overall df for comparison

test_mfq_mutate <- cbind(test_mfq_pre, test_mfq_post)

```

```{r testing function to setup future mutation checks, include=FALSE}

# Example usage:
# Assuming `data_post_named` is the pre-mutation dataframe
# and `data_mfq_mutated` is the post-mutation dataframe,
# and you want to compare the value counts of "mfq_1":

test_mutate <- prep_mutation_check(data_post_named, data_mfq_mutated, "mfq_1") # same output as test_mfq_mutate above

## will use the prep_mutation_check() function to set up mutation checks moving forward

```

Checking mfq response counts are identical from pre to post mutation

```{r mfq response counts validation, validate = TRUE}

test_mutate %>%
  pointblank::col_vals_equal(vars(count_pre), 
                             vars(count_post)) # passes if response counts are identical pre to post mutation

```

Calculating mfq composite scores

```{r calculating mfq scores}

## Calculating composite mfq scores

data_mfq_scored <- data_mfq_mutated %>%
  mutate(mfq_total = rowSums(across(starts_with("mfq_") & -contains("tim")), na.rm = TRUE)) %>%
  mutate(mfq_total = case_when(
    is.na(mfq_1) & is.na(mfq_2) & is.na(mfq_3) & is.na(mfq_4) & is.na(mfq_5) &
    is.na(mfq_6) & is.na(mfq_7) & is.na(mfq_8) & is.na(mfq_9) & is.na(mfq_10) &
    is.na(mfq_11) & is.na(mfq_12) & is.na(mfq_13)
     ~ NA_real_,
    TRUE ~ mfq_total
  ))

```

```{r checking mfq score setup, include = FALSE}

## checking composite calculation was correct

# manually calculating mfq totals to compare sum methods

test_mfq_calc <- data_mfq_scored %>%
  select(response_id, starts_with("mfq_") & -contains("tim")) %>%
  mutate(manual_mfq_total = mfq_1 + mfq_2 + mfq_3 + mfq_4 + mfq_5 +
                            mfq_6 + mfq_7 + mfq_8 + mfq_9 + mfq_10 +
                            mfq_11 + mfq_12 + mfq_13) %>%
  select(mfq_total, manual_mfq_total)

```

Checking mfq totals are identical across sum methods

```{r mfq scoring validation, validate = TRUE}

test_mfq_calc %>%
  pointblank::col_vals_equal(vars(mfq_total), 
                             vars(manual_mfq_total), 
                             na_pass = TRUE) # passes if mfq totals are identical across sum methods

```

### Hopelessness

Calculating hopelessness using the 4-item Beck's Hopelessness Scale (measured pre and post SSI). Row mean scores range from 0 to 3, where 0 = absolutely disagree, 1 = somewhat disagree, 2 = somewhat agree, 3 = absolutely agree on each item. Source: https://doi.org/10.1007/s10597-013-9619-1

```{r pulling bhs values, include=FALSE}

## Pulling mfq variables

vars_bhs <- data_mfq_scored %>%
  select(contains("bhs")) %>%
  head() %>%
  print()

table(vars_bhs$pre_bhs_4)

```

Mutating bhs character values to numeric values

```{r mutating bhs values}

## Mutating character values to numeric across multiple bhs columns 

data_bhs_mutated <- data_mfq_scored %>%
  mutate(across(contains("bhs_") & -contains("tim"), 
                ~ case_when(
                    . == "Kesinlikle katılmıyorum" ~ 0,
                    . == "Kısmen katılmıyorum" ~ 1,
                    . == "Kısmen katılıyorum" ~ 2,
                    . == "Kesinlikle katılıyorum" ~ 3,
                    TRUE ~ NA_real_ # Handle other cases with NA
                )))

```

Checking bhs variables are now numeric class

```{r bhs numeric class check}

## Checking that all bhs vars are numeric class

data_bhs_mutated %>%
  select(contains("bhs")) %>%
  summarise(across(everything(), ~ class(.x)[1])) %>%
  tidyr::pivot_longer(cols = everything(), names_to = "variable", values_to = "class")


```

```{r using function to setup bhs mutation check, include=FALSE}

# `data_mfq_scored` is the pre-mutation dataframe
# and `data_bhs_mutated` is the post-mutation dataframe,
# and we want to compare the value counts of "pre_bhs_1":

test_mutate <- prep_mutation_check(data_mfq_scored, data_bhs_mutated, "pre_bhs_1")

```

Checking bhs response counts are identical from pre to post mutation

```{r bhs response counts validation, validate = TRUE}

test_mutate %>%
  pointblank::col_vals_equal(vars(count_pre), 
                             vars(count_post)) # passes if response counts are identical pre to post mutation

# test_pre_bhs <- data_bhs_mutated %>%
#   filter(pre_bhs_1 == 3) %>%
#   select(response_id)
# 
# test_post_bhs <- data_mfq_scored %>%
#   filter(pre_bhs_1 == "Kesinlikle katılıyorum") %>%
#   select(response_id)
# 
# identical(test_pre_bhs, test_post_bhs) # also passes visual specific count inspection

```

Calculating bhs row means

```{r calculating bhs scores}

## Calculating composite scores

data_bhs_scored <- data_bhs_mutated %>%
  mutate(pre_bhs_mean = rowMeans(across(starts_with("pre_bhs_") & -contains("tim")), na.rm = TRUE),
         post_bhs_mean = rowMeans(across(starts_with("post_bhs_") & -contains("tim")), na.rm = TRUE)) %>%
  mutate(across(contains("bhs_"), ~ case_when(
                                  is.nan(.) ~ NA_real_, # Replace NaN with NA
                                  TRUE ~ .))) # Keep other values unchanged

```

```{r checking bhs score setup, include = FALSE}

## checking composite calculation was correct

# manually calculating pre means to compare methods

test_pre_bhs_calc <- data_bhs_scored %>%
  select(response_id, starts_with("pre_bhs_") & -contains("tim")) %>%
  mutate(pre_manual_mean = rowMeans(select(., pre_bhs_1, pre_bhs_2, pre_bhs_3, pre_bhs_4), na.rm = TRUE)) %>%
  mutate(across(contains("manual"), ~ case_when(is.nan(.) ~ NA_real_, # Replace NaN with NA
                                        TRUE ~ .))) %>% # Keep other values unchanged
  select(pre_bhs_mean, pre_manual_mean)

# manually calculating post means to compare methods

test_post_bhs_calc <- data_bhs_scored %>%
  select(response_id, starts_with("post_bhs_") & -contains("tim")) %>%
  mutate(post_manual_mean = rowMeans(select(., post_bhs_1, post_bhs_2, post_bhs_3, post_bhs_4), na.rm = TRUE)) %>%
  mutate(across(contains("manual"), ~ case_when(is.nan(.) ~ NA_real_, # Replace NaN with NA
                                        TRUE ~ .))) %>% # Keep other values unchanged
  select(post_bhs_mean, post_manual_mean)

```

Checking pre and post bhs means are identical across averaging methods

```{r pre bhs scoring validation, validate = TRUE}

test_pre_bhs_calc %>%
  pointblank::col_vals_equal(vars(pre_bhs_mean), 
                             vars(pre_manual_mean), 
                             na_pass = TRUE) # passes if means are identical across methods

```

```{r post bhs scoring validation, validate = TRUE}

test_post_bhs_calc %>%
  pointblank::col_vals_equal(vars(post_bhs_mean), 
                             vars(post_manual_mean), 
                             na_pass = TRUE) # passes if means are identical across methods

```

### Hope (Pathways)

Calculating hope using the 3-item pathways subscale from the State Hope Scale (measured pre and post SSI). Row mean scores range from 1 to 8, where 1 = definitely false, and 8 = definitely true. Source: https://doi.org/10.1037/0022-3514.70.2.321

```{r pulling shs values, include=FALSE}

## Pulling shs variables

vars_shs <- data_bhs_scored %>%
  select(contains("shs")) %>%
  head() %>%
  print()

```

Mutating shs character values to numeric values

```{r mutating shs values}

## Mutating character values to numeric across multiple shs columns 

data_shs_mutated <- data_bhs_scored %>%
  mutate(across(contains("shs_") & -contains("tim"), 
                ~ case_when(
                    . == "Kesinlikle Yanlış" ~ 1,
                    . == "Büyük Ölçüde Yanlış" ~ 2,
                    . == "Kısmen Yanlış" ~ 3,
                    . == "Biraz Yanlış" ~ 4,
                    . == "Biraz Doğru" ~ 5,
                    . == "Kısmen Doğru" ~ 6,
                    . == "Büyük Ölçüde Doğru" ~ 7,
                    . == "Kesinlikle Doğru" ~ 8,
                    TRUE ~ NA_real_ # Handle other cases with NA
                )))
 
```

Checking shs variables are now numeric class

```{r shs numeric class check}

## Checking that all shs vars are numeric class

data_shs_mutated %>%
  select(contains("shs")) %>%
  summarise(across(everything(), ~ class(.x)[1])) %>%
  tidyr::pivot_longer(cols = everything(), names_to = "variable", values_to = "class")


```

```{r using function to setup shs mutation check, include=FALSE}

# `data_bhs_scored` is the pre-mutation dataframe
# and `data_shs_mutated` is the post-mutation dataframe,
# and we want to compare the value counts of "pre_shs_1":

test_mutate <- prep_mutation_check(data_bhs_scored, data_shs_mutated, "pre_shs_1")

```

Checking shs response counts are identical from pre to post mutation

```{r shs response counts validation, validate = TRUE}

test_mutate %>%
  pointblank::col_vals_equal(vars(count_pre), 
                             vars(count_post)) # passes if response counts are identical pre to post mutation

```

Calculating shs row means

```{r calculating shs scores}

## Calculating composite scores

data_shs_scored <- data_shs_mutated %>%
  mutate(pre_shs_mean = rowMeans(across(starts_with("pre_shs_") & -contains("tim")), na.rm = TRUE),
         post_shs_mean = rowMeans(across(starts_with("post_shs_") & -contains("tim")), na.rm = TRUE)) %>%
  mutate(across(contains("shs_"), ~ case_when(
                                  is.nan(.) ~ NA_real_, # Replace NaN with NA
                                  TRUE ~ .))) # Keep other values unchanged

```

```{r checking shs score setup, include = FALSE}

## checking composite calculation was correct

# manually calculating pre means to compare methods

test_pre_shs_calc <- data_shs_scored %>%
  select(response_id, starts_with("pre_shs_") & -contains("tim")) %>%
  mutate(pre_manual_mean = rowMeans(select(., pre_shs_1, pre_shs_2, pre_shs_3), na.rm = TRUE)) %>%
  mutate(across(contains("manual"), ~ case_when(is.nan(.) ~ NA_real_, # Replace NaN with NA
                                        TRUE ~ .))) %>% # Keep other values unchanged
  select(pre_shs_mean, pre_manual_mean)

# manually calculating post means to compare methods

test_post_shs_calc <- data_shs_scored %>%
  select(response_id, starts_with("post_shs_") & -contains("tim")) %>%
  mutate(post_manual_mean = rowMeans(select(., post_shs_1, post_shs_2, post_shs_3), na.rm = TRUE)) %>%
  mutate(across(contains("manual"), ~ case_when(is.nan(.) ~ NA_real_, # Replace NaN with NA
                                        TRUE ~ .))) %>% # Keep other values unchanged
  select(post_shs_mean, post_manual_mean)

```

Checking pre and post shs means are identical across averaging methods

```{r pre shs scoring validation, validate = TRUE}

test_pre_shs_calc %>%
  pointblank::col_vals_equal(vars(pre_shs_mean), 
                             vars(pre_manual_mean), 
                             na_pass = TRUE) # passes if means are identical across methods

```

```{r post shs scoring validation, validate = TRUE}

test_post_shs_calc %>%
  pointblank::col_vals_equal(vars(post_shs_mean), 
                             vars(post_manual_mean), 
                             na_pass = TRUE) # passes if means are identical across methods

```

### Self-Hate

Calculating self-hate using a 3-item version of the Self-Hate Scale (measured pre and post SSI). Row mean scores range from 1 to 7, where 1 = not at all true for me, and 7 = very true for me. Source (abbreviated scale): https://mental.jmir.org/2020/6/e20513. Source (full scale): https://doi.org/10.1016/j.jad.2018.11.047 

NOTE: While we've used 6-point Likert scales for items in our English data and in the original YES data, we use the full 7-point Likert scale items here. 

```{r pulling selfhate values, include=FALSE}

## Pulling selfhate variables

vars_hate <- data_shs_scored %>%
  select(contains("self_hate")) %>%
  head() %>%
  print()

```

Mutating hate character values to numeric values

```{r mutating hate values}

## Converting number character values from sliding scale to numeric values

data_hate_mutated <- data_shs_scored %>%
  mutate(across(contains("self_hate"), 
                ~ case_when(
                  . == "Benim için hiç uygun değil" ~ "1",
                  . == "2" ~ "2",
                  . == "3" ~ "3",
                  . == "4" ~ "4",
                  . == "5" ~ "5",
                  . == "6" ~ "6",
                  . == "Benim için tamamen uygun" ~ "7",
                  TRUE ~ NA_character_))) %>%
  mutate(across(contains("self_hate"), 
                ~ as.numeric(.)))

```

Checking hate variables are now numeric class

```{r hate numeric class check}

## Checking that all self-hate vars are numeric class

data_hate_mutated %>%
  select(contains("self_hate")) %>%
  summarise(across(everything(), ~ class(.x)[1])) %>%
  tidyr::pivot_longer(cols = everything(), names_to = "variable", values_to = "class")


```

Calculating hate row means

```{r calculating hate scores}

## Calculating composite scores

data_hate_scored <- data_hate_mutated %>%
  mutate(pre_self_hate_mean = rowMeans(across(starts_with("pre_self_hate_") & -contains("tim")), na.rm = TRUE),
         post_self_hate_mean = rowMeans(across(starts_with("post_self_hate_") & -contains("tim")), na.rm = TRUE)) %>%
  mutate(across(contains("self_hate_"), ~ case_when(
                                  is.nan(.) ~ NA_real_, # Replace NaN with NA
                                  TRUE ~ .))) # Keep other values unchanged

```

```{r checking hate score setup, include = FALSE}

## checking composite calculation was correct

# manually calculating pre means to compare methods

test_pre_hate_calc <- data_hate_scored %>%
  select(response_id, starts_with("pre_self_hate_") & -contains("tim")) %>%
  mutate(pre_manual_mean = rowMeans(select(., pre_self_hate_1, pre_self_hate_2, pre_self_hate_3), na.rm = TRUE)) %>%
  mutate(across(contains("manual"), ~ case_when(is.nan(.) ~ NA_real_, # Replace NaN with NA
                                        TRUE ~ .))) %>% # Keep other values unchanged
  select(pre_self_hate_mean, pre_manual_mean)

# manually calculating post means to compare methods

test_post_hate_calc <- data_hate_scored %>%
  select(response_id, starts_with("post_self_hate_") & -contains("tim")) %>%
  mutate(post_manual_mean = rowMeans(select(., post_self_hate_1, post_self_hate_2, post_self_hate_3), na.rm = TRUE)) %>%
  mutate(across(contains("manual"), ~ case_when(is.nan(.) ~ NA_real_, # Replace NaN with NA
                                        TRUE ~ .))) %>% # Keep other values unchanged
  select(post_self_hate_mean, post_manual_mean)

```

Checking pre and post hate means are identical across averaging methods

```{r pre hate scoring validation, validate = TRUE}

test_pre_hate_calc %>%
  pointblank::col_vals_equal(vars(pre_self_hate_mean), 
                             vars(pre_manual_mean), 
                             na_pass = TRUE) # passes if means are identical across methods

```

```{r post hate scoring validation, validate = TRUE}

test_post_hate_calc %>%
  pointblank::col_vals_equal(vars(post_self_hate_mean), 
                             vars(post_manual_mean), 
                             na_pass = TRUE) # passes if means are identical across methods

```

### PFS

Calculating program feedback scores using the Program feedback scale (measured post SSI). Row mean scores range from 1 to 5, where 1 = really disagree, and 5 = really agree. Source: https://mental.jmir.org/2020/6/e20513

```{r pulling pfs values, include=FALSE}

## Pulling pfs variables

# vars_pfs <- data_hate_scored %>%
#   select(contains("pfs") & -contains("like") &
#            -contains("change") &
#            -contains("more") &
#            -contains("tim")) %>%
#   head() %>%
#   print()

```

Mutating pfs character values to numeric values

```{r mutating pfs values}

## Mutating character values to numeric across multiple pfs columns 

data_pfs_mutated <- data_hate_scored %>%
  mutate(across(contains("pfs_") & -contains("like") & 
               -contains("change") &
               -contains("more") &
               -contains("tim"),
                ~ case_when(
                    . == "Hiç katılmıyorum" ~ 1,
                    . == "Katılmıyorum" ~ 2,
                    . == "Ne katılıyorum, ne katılmıyorum" ~ 3,
                    . == "Katılıyorum" ~ 4,
                    . == "Tamamen katılıyorum" ~ 5,
                    TRUE ~ NA_real_ # Handle other cases with NA
                )))
 
```

Checking pfs variables are now numeric class

```{r pfs numeric class check}

## Checking that all pfs vars are numeric class

data_pfs_mutated %>%
  select(contains("pfs") & 
         -contains("like") & 
         -contains("change") &
         -contains("more") &
         -contains("tim"),) %>%
  summarise(across(everything(), ~ class(.x)[1])) %>%
  tidyr::pivot_longer(cols = everything(), names_to = "variable", values_to = "class")


```

```{r using function to setup pfs mutation check, include=FALSE}

# `data_hate_scored` is the pre-mutation dataframe
# and `data_pfs_mutated` is the post-mutation dataframe,
# and we want to compare the value counts of "pfs_1":

test_mutate <- prep_mutation_check(data_hate_scored, data_pfs_mutated, "pfs_1")

```

Checking pfs response counts are identical from pre to post mutation

```{r pfs response counts validation, validate = TRUE}

test_mutate %>%
  pointblank::col_vals_equal(vars(count_pre), 
                             vars(count_post)) # passes if response counts are identical pre to post mutation

```

Calculating pfs row means

```{r calculating pfs scores}

## Calculating composite scores

data_pfs_scored <- data_pfs_mutated %>%
  # Calculate the row means of relevant pfs_ columns, excluding specific patterns
  mutate(pfs_mean = rowMeans(across(contains("pfs_") & 
                                    !contains("like") & 
                                    !contains("change") &
                                    !contains("more") &
                                    !contains("tim")), na.rm = TRUE)) %>%
  # Replace NaN values with NA_real_ across relevant pfs columns
  mutate(across(contains("pfs_") &
                !contains("like") & 
                !contains("change") &
                !contains("more") &
                !contains("tim"),
                ~ case_when(is.nan(.x) ~ NA_real_,  # Replace NaN with NA
                            TRUE ~ .x))) # Keep other values unchanged

```

```{r checking pfs score setup, include = FALSE}

## checking composite calculation was correct

# manually calculating means to compare methods

test_pfs_calc <- data_pfs_scored %>%
  select(response_id, starts_with("pfs") & 
                                    !contains("like") & 
                                    !contains("change") &
                                    !contains("more") &
                                    !contains("tim")) %>%
  mutate(manual_mean = rowMeans(select(., pfs_1, pfs_2, pfs_3, pfs_4, pfs_5, pfs_6, pfs_7), na.rm = TRUE)) %>%
  mutate(across(contains("manual"), ~ case_when(is.nan(.) ~ NA_real_, # Replace NaN with NA
                                        TRUE ~ .))) %>% # Keep other values unchanged
  select(pfs_mean, manual_mean)

```

Checking pfs mean is identical across averaging methods

```{r pfs scoring validation, validate = TRUE}

test_pfs_calc %>%
  pointblank::col_vals_equal(vars(pfs_mean), 
                             vars(manual_mean), 
                             na_pass = TRUE) # passes if means are identical across averaging methods

```

## Reorganizing Columns

Dropping unnecessary columns

```{r dropping cols}

data_col_dropped <- data_pfs_scored %>%
  select(-status, -contains("duration"), -contains("recipient"), -contains("external"), -contains("distribution"), 
         -recorded_date, -user_language) %>% # dropping unnecessary qualtrics metadata
  select(-sc0, -sc1, -sc2) %>% # dropping scale scores calculated within qualtrics
  select(-questionnaires) # dropping col redundant with volunteer col

```

Adding cols to document language and data source

```{r adding cols}

data_col_added <- data_col_dropped %>%
  mutate(language = "turkish",
         data_id = "d07")

# glimpse(data_col_added)

```

Reorganizing column order to place important columns first, and then to follow chronological order

```{r reorganizing col order}

## Reorganizing cols to prioritize important cols first and then chronological order

data_reorged <- data_col_added %>%
  select(response_id, language, data_id, contains("date"), ip_address, starts_with("location"),
         perc_progress, finished, q_recaptcha_score, ssi_choice,
         starts_with("volunteer_") & !contains("tim"),
         dem_hear_proj, dem_why_interest, dem_receive_support, dem_why_no_support, dem_age, dem_sex,
         starts_with("dem_gender_") & !contains("tim"),
         starts_with("dem_sex_orientation") & !contains("tim"), 
         starts_with("dem_race") & !contains("tim"),
         starts_with("ssi_prev"), starts_with("ssi_times"), 
         starts_with("mfq_") & !contains("tim"), 
         starts_with("pre_bhs_") & !contains("tim"), 
         starts_with("pre_shs_") & !contains("tim"), 
         starts_with("pre_self_hate_") & !contains("tim"),
         starts_with("pre_perceived_control") & !contains("tim"),
         ssi_started,
         starts_with("abc") & !contains("tim"),
         starts_with("proj_care") & !contains("tim"),
         starts_with("proj_pers") & !contains("tim"),
         starts_with("proj_rise") & !contains("tim"),
         ssi_completed,
         starts_with("advice") & !contains("tim"),
         starts_with("pfs_") & !contains("tim"),
         starts_with("perc_change_") & !contains("tim"),
         starts_with("post_bhs_") & !contains("tim"),
         starts_with("post_shs_") & !contains("tim"),
         starts_with("post_self_hate_") & !contains("tim"),
         starts_with("post_perceived_control") & !contains("tim"),
         contains("tim") & starts_with("intro"),
         contains("tim") & starts_with("volunteer"),
         contains("tim") & starts_with("dem"),
         contains("tim") & starts_with("ssi"),
         contains("tim") & starts_with("mfq"),
         contains("tim") & starts_with("pre_"),
         contains("tim") & starts_with("abc"),
         contains("tim") & starts_with("proj_care"),
         contains("tim") & starts_with("proj_rise"),
         contains("tim") & starts_with("proj_pers"),
         contains("tim") & starts_with("advice_"),
         contains("tim") & starts_with("pfs_"),
         contains("tim") & starts_with("perc_change_"),
         contains("tim") & starts_with("post_"))

```

```{r checking col order setup, include=FALSE}

## creating data frames of names pre and post reorg for comparison

pre_reorg_names <- data_col_added %>%
  names() # names pre

post_reorg_names <- data_reorged %>%
  names() # names post

# Create tibbles for easier manipulation

pre_reorg_df <- tibble(names = pre_reorg_names)
post_reorg_df <- tibble(names = post_reorg_names)

# Get the unique names (names not in both data frames)

test_reorg <- bind_rows(pre_reorg_df, post_reorg_df) %>%
  group_by(names) %>%
  filter(n() == 1)
         
```

Checking that reorganizing happened correctly (i.e., no columns dropped) 

```{r checking reorg did not lose cols 2}

## Checking if df with any columns unique to pre-reorg data is empty (i.e., no columns dropped)

n_cols_dropped <- test_reorg %>%
  nrow() == 0

n_cols_dropped # true if no columns were dropped by reorg

```

## Converting Classes

Converting variable classes

```{r mutating var classes}

data_classes_mutated <- data_reorged %>%
  select(-ip_address, -starts_with("location")) %>%
  mutate(language = as.factor(language),
         data_id = as.factor(data_id),
         perc_progress = as.numeric(perc_progress),
         finished = as.factor(finished),
         volunteer_questionnaires_yn = as.factor(volunteer_questionnaires_yn),
         pre_perceived_control = as.numeric(pre_perceived_control),
         post_perceived_control = as.numeric(post_perceived_control),
         ssi_choice = as.factor(ssi_choice),
         ssi_started = as.factor(ssi_started),
         ssi_completed = as.factor(ssi_completed)) %>%
  mutate(across(contains("date"), as.Date)) %>%
  mutate(across(matches("^dem_.*") & !matches("why|hear|text"), as.factor)) %>%
  mutate(across(contains("times_completed"), as.numeric)) %>%
  mutate(across(contains("ssi_prev_completed"), as.factor)) %>%
  mutate(across(contains("_q_") & ends_with("_num"), as.numeric)) %>% # ssi numeric questions
  mutate(across(contains("_q_") & -ends_with("_text") & -ends_with("_num") &
                -ends_with("_rank") & -ends_with("_group"), as.factor)) %>% # ssi qual questions
  mutate(across(contains("commit"), as.factor)) %>%
  mutate(across(contains("perc_change"), as.factor)) %>%
  mutate(across(contains("_tim_"), as.numeric))

```

# Saving Data

```{r saving cleaned data}

data <- data_classes_mutated

# Define the folder and file path

folder_path <- here("data", "clean")  # Path to my desired folder
file_name <- "yes_data_d07_turkish_clean.Rds"
full_path <- file.path(folder_path, file_name)

# Save the dataframe to the .csv file

saveRDS(data, full_path)

```
